---
title: "GT & HL Incidents by Installation"
subtitle: "Analyzing Options & Alternatives"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  ioslides_presentation:
    widescreen: true
    logo: PST_logo.png
    css: styles.css
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE, echo = FALSE}
rm(list = ls())
```

```{r setup, include=FALSE}
library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(sp)
library(rgeos)
library(tufte)
library(rgdal)
library(ggspatial)
library(mapboxapi)

knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
options(knitr.table.format = "latex")
wdt = 2560
hgt = wdt/1.62

datadir <- "https://raw.githubusercontent.com/jmceager/pst/main/phmsa-clean/data/"
capt = "Source: PHMSA Incident and Mileage Data (2010-2021)"
tag = "PST 2022"
```

```{r theme_pst}
library(showtext)
#> Loading required package: sysfonts
#> Loading required package: showtextdb
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()
theme_pst <- function(baseSize=8) {
  (theme(
                       axis.line.x = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.line.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.ticks.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       text = element_text(family = pstFont),
                       axis.text = element_text(colour = "#8C9394",
                                                size = baseSize * .8),
                       axis.title = element_text(colour ="black"),
                       panel.grid.minor.y = element_line(linetype = "dotted", 
                                                         colour = "#C4C8C6"),
                       panel.grid.major.y = element_line(colour = "#394C56", 
                                                         linetype = "dotted"),
                       panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(),
                       panel.background = element_blank(),
                       panel.border = element_blank(),
                       strip.background = element_rect(linetype = 0),
                       strip.text = element_blank(),
                       strip.text.x = element_blank(),
                       strip.text.y = element_blank(),
                       legend.text = element_text(colour = "black",
                                                  size = baseSize * .8),
                       legend.background = element_rect(fill = NULL, 
                                                        color = "#182125"),
                       legend.title = element_text( face = "bold", 
                                                    colour = "black",
                                                    size = baseSize),
                       legend.position = "right",
                       legend.key = element_blank(),
                       #legend.background = element_blank(),
                       plot.tag.position = "topright",
                       plot.tag = element_text(size = baseSize *.66,
                                               colour = "#8C9394"),
                       plot.background = element_blank(),
                       plot.title = element_text(face = "bold", 
                                                 colour = "black",
                                                 size = baseSize*1.2),
                       plot.subtitle = element_text(colour = "#8C9394",
                                                    size = baseSize)
                     )
  )
}
```

```{r nice colors}
#brewer pal
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]

#pst 
darkBlue <- "#003E59"
midBlue <- "#00546F"
bluGreen <- "#459197"
liteGreen <- "#61A893"
```

```{r loading data}
# no gd because of too many mitigating factors wrt materials, etc. 
## mileage
gt.mi <- read.csv(paste0(datadir,"raw/GT_MilesDecadeAge.csv"))
hl.mi <- read.csv(paste0(datadir,"raw/HL_MilesDecadeAge.csv"))

## incidents
gt.inc <- read_csv(paste0(datadir,"clean/all_inc.csv")) %>% filter(grepl("GT", SYS)) %>%
  mutate(STATE = if_else(grepl("Outer Continental Shelf", cleanLoc), "OCSG", STATE))
hl.inc <- read_csv(paste0(datadir,"clean/all_inc.csv")) %>% filter(grepl("HL", SYS)) %>%
  mutate(STATE = if_else(grepl("Outer Continental Shelf", cleanLoc), 
                         if_else(STATE == "CA", "OCSP", "OCSG"),
                         STATE))
```

```{r mile normalization}
# grouping unknown & pre40s
gtMileNorm <- gt.mi %>% 
  select(Calendar.Year, starts_with("X"), starts_with("Pre"))%>%
  rename(X1930 = Pre.1940.or.Unknown) %>%
  filter(Calendar.Year >= 2010)%>%
  group_by(Calendar.Year)%>%
  summarise(across(c(1:10),~sum(.,na.rm = T))) %>%
  pivot_longer(cols = starts_with("X"))%>%
  mutate(name = parse_number(str_sub(name, 2,5)))

#separating unknown & pre40s
gtMileKnown <- gt.mi %>% 
  rename(X1930 = Trans.pre.1940.Onshore.Miles,
         X1920 = Trans.Unknown.by.Dec.Onshore.Miles) %>%
  select(Calendar.Year, starts_with("X"))%>%
  filter(Calendar.Year >= 2010)%>%
  group_by(Calendar.Year)%>%
  summarise(across(c(1:11),~sum(.,na.rm = T))) %>%
  pivot_longer(cols = starts_with("X"))%>%
  mutate(name = parse_number(str_sub(name, 2,5)))

#liquids 
hlMileNorm <- hl.mi %>%
  rename(X1910 = Pre.20.or.Unknown) %>%
  select(Calendar.Year, starts_with("X")) %>%
  filter(Calendar.Year >= 2010)%>%
  group_by(Calendar.Year)%>%
  summarise(across(c(1:11),~sum(.,na.rm = T)))%>%
  pivot_longer(cols = starts_with("X"))%>%
  mutate(name = parse_number(str_sub(name, 2,5)))
```

## Guiding Questions 

- Who is the audience for these plots and discussions?  
    - What are we presenting externally vs using internally?  
    - What level of depth is required?
- How defensible or credible are some of the choices we are making?  
- What is the story behind the data, inclusive of all choices?

# Gas Transmission

## Base Plot

```{r base plot, include = TRUE}
gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>% 
  mutate(s.ipm = if_else(INSTALLATION_DECADE == 2020, s.ipm*5, s.ipm))%>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm))+ 
  geom_bar(stat = "identity", fill = midBlue)+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.7)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1925,2025),
                     breaks = seq(1930, 2010,20),
                     labels = c("Pre-40s\n& Unkown",seq(1950,2010,20)))+
  scale_y_continuous(expand = expansion(mult = c(0,.1), add=0))+
  scale_fill_brewer(palette = "Paired", direction = -1)+
  theme(legend.position = c(.7,.7))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents of Onshore Gas Transmission Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")
```

## Offshore Pipeline or Compressor Stations

```{r add offshore pipe or compressor, include = TRUE}
## add offshore incidents 
gt.offshore <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" |
           SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND"  ,
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ Offshore")

##add compressors, but NOT offshore
gt.compressor <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" |
         SYSTEM_PART_INVOLVED == "ONSHORE COMPRESSOR STATION EQUIPMENT AND PIPING"     ,
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ Compressor")



gt.base <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>% 
  mutate(version = "Base")

rbind(gt.base, gt.compressor, gt.offshore)%>%
  mutate(version = factor(version, levels = c("Base", "+ Offshore","+ Compressor"))) %>%
  mutate(s.ipm = if_else(INSTALLATION_DECADE == 2020, s.ipm*5, s.ipm))%>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm, 
             fill = version, group = version))+ 
  geom_bar(stat = "identity",position = "dodge")+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.7)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1925,2025),
                     breaks = seq(1930, 2010,20),
                     labels = c("Pre-40s\n& Unkown",seq(1950,2010,20)))+
  scale_y_continuous(expand = expansion(mult = c(0,.1), add=0))+
  scale_fill_manual(name = "Parts", values = c(darkBlue, midBlue, bluGreen))+
  theme(legend.position = c(.7,.7))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents on Gas Transmission Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")
```

## Why not both?

```{r offshore and comp, include = TRUE}
gt.offcomp <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" |
           SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND" |
           SYSTEM_PART_INVOLVED == "ONSHORE COMPRESSOR STATION EQUIPMENT AND PIPING",
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ OffShr + Comp")

rbind(gt.base, gt.offcomp)%>%
  mutate(version = factor(version, levels = c("Base", "+ OffShr + Comp"))) %>%
  mutate(s.ipm = if_else(INSTALLATION_DECADE == 2020, s.ipm*5, s.ipm))%>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm, 
             fill = version, group = version))+ 
  geom_bar(stat = "identity",position = "dodge")+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.7)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1925,2025),
                     breaks = seq(1930, 2010,20),
                     labels = c("Pre-40s\n& Unkown",seq(1950,2010,20)))+
  scale_y_continuous(expand = expansion(mult = c(0,.1), add=0))+
  scale_fill_manual(name = "Version", values = c(darkBlue, bluGreen))+
  theme(legend.position = c(.7,.7),
        panel.border = element_blank())+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents on Gas Transmission Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")

```

## 1970s Offshore | Anomaly or Expected? 


```{r}
gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND",
         IYEAR < 2022)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>% view()
```

```{r need to extract percent OCS}
gt70sinc <- gt.inc %>% 
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND",
         IYEAR < 2022,
         between(INSTALLATION_YEAR, 1970, 1980))%>%
  mutate(type = if_else(grepl("OCS", STATE),
                        "OCS",
                        "IS"))%>%
  group_by(type)%>%
  summarise(n = n())

per70sinc <- round((gt70sinc[[2,"n"]] / sum(gt70sinc$n) )*100, 1)

gtAllInc <- gt.inc %>% 
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND",
         IYEAR < 2022,
         !between(INSTALLATION_YEAR, 1970,1980))%>%
  mutate(type = if_else(grepl("OCS", STATE),
                        "OCS",
                        "IS"))%>%
  group_by(type)%>%
  summarise(n = n())

perAllInc <- round((gtAllInc[[2,"n"]] / sum(gtAllInc$n) )*100, 1)
```

-  %`r per70sinc` of 1970s-Installed incidents were on the Outer Continental Shelf  
    -   Compared to %`r perAllInc` in all other installation decades   
-  A vast majority of OCS Mileage in use as of 2021 appears to be built post-1960     
    -   In-state pipelines offer a more balanced proportion, as we'll see on the next plot  
 



```{r OCS vs rest mileage, include = TRUE}
gt.mi %>%
  filter(Calendar.Year == 2021)%>%
  mutate(type = if_else(grepl("OCS", State.Abbreviation),
                        "OCS",
                        "IS")) %>%
  rename(X1930 = Pre.1940.or.Unknown)%>%
  select(type, starts_with("X"))%>%
  group_by(type)%>%
  summarise(across(starts_with("X"), ~ sum(.x, na.rm = T)))%>%
  mutate(total = rowSums(across(where(is.numeric))),
         across(starts_with("X"), ~ .x/total))%>%
  select(!total)%>%
  pivot_longer(!type, names_to = "Decade", values_to = "Percent")%>%
  mutate(Decade = str_sub(Decade,2,5),
         Decade = as.numeric(Decade))%>%
  ggplot(aes(x = Decade, y = Percent, group = type, fill = type))+
  geom_bar(stat = "identity", position = "dodge")+ 
  geom_text(aes(label = round(Percent*100,1), family = pstFont), 
            vjust = -1, colour = "gray15", position = position_dodge(width = 9.3), size = 2.7)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0,.1)),
                     name = "Percent Total Mileage")+
  scale_fill_manual(name = "Geography",
                    values = c(darkBlue, bluGreen))+
  scale_x_continuous(limits = c(1925,2025),
                     breaks = seq(1930, 2010,20),
                     labels = c("Pre-40s\n& Unkown",seq(1950,2010,20)))+
  theme_pst(baseSize = 12)+
  theme(legend.position = c(.8,.7))+
  labs(title = "Mileage by Pipeline Geography",
       subtitle = "Percent of Total Mileage, 2021",
       caption = capt,
       tag = tag)

```


## Unknown Mileage | A Refresh

```{r refresh unknown, include = TRUE}

gt.unknown.group <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,                 
                                          na = c("NA","UNKNOWN","N/A")),    
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),           
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",   
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR !=2022,
         IYEAR > 2011)%>%
  left_join(gtMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T)%>%                                                        
  #filter(!is.na(value))%>%                                                  
  group_by(INSTALLATION_DECADE, IYEAR)%>%                                       
  summarise(n = n(),                                                         
            miles = mean(value),                                             
            ipm = n/(miles/1000)) %>%                                        
  group_by(INSTALLATION_DECADE)%>%            
  mutate(count = if_else(INSTALLATION_DECADE == 2020, 
                         as.integer(2),
                         as.integer(10)),
         count.u = if_else(INSTALLATION_DECADE == 2020, 
                         as.integer(5),
                         as.integer(10)))%>%
  summarise(total = sum(n),   
            iyear = IYEAR, 
            w.ipm = ipm * (n/total),   
            count = count,
            s.ipm = round(sum(w.ipm)*count,1),
            u.ipm = round(mean(ipm)*count.u,1))%>%                            
  distinct(INSTALLATION_DECADE,.keep_all = T)%>%
  ungroup()%>%
  add_row(INSTALLATION_DECADE = 1920, u.ipm = 0.0)%>%
  mutate(version = "Grouped")
  
  
gt.unknown.split <- gt.inc %>%
  #drop_na(INSTALLATION_YEAR)%>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,na = c("NA","UNKNOWN")),             
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE),
         INSTALLATION_DECADE = replace_na(INSTALLATION_DECADE, 1920))%>%
  filter(SIGNIFICANT == "YES",   
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR !=2022,
         IYEAR > 2011)%>%
  left_join(gtMileKnown,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T)%>%                                                        
  #filter(!is.na(value))%>%                                                  
  group_by(INSTALLATION_DECADE, IYEAR)%>%                                       
  summarise(n = n(),                                                         
            miles = mean(value),                                             
            ipm = n/(miles/1000)) %>%                                         
  group_by(INSTALLATION_DECADE)%>%            
  mutate(count = if_else(INSTALLATION_DECADE == 2020, 
                         as.integer(2),
                         as.integer(10)))%>%
  mutate(count = if_else(INSTALLATION_DECADE == 2020, 
                         as.integer(2),
                         as.integer(10)),
         count.u = if_else(INSTALLATION_DECADE == 2020, 
                         as.integer(5),
                         as.integer(10)))%>%
  summarise(total = sum(n),   
            iyear = IYEAR, 
            w.ipm = ipm * (n/total),   
            count = count,
            s.ipm = round(sum(w.ipm)*count,1),
            u.ipm = round(mean(ipm)*count.u,1))%>%                            
  distinct(INSTALLATION_DECADE,.keep_all = T)%>%
  mutate(version = "Separate")

rbind(gt.unknown.split, gt.unknown.group)%>%   
  ggplot(aes(x = INSTALLATION_DECADE, y = u.ipm, fill = version, group = version))+ 
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = u.ipm, family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.5)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1915,2025),
                     breaks = seq(1920, 2020,10),
                     labels = c("Unknown", "Pre-40s", 
                                paste0("'",seq(40,90,10),"s"),
                                "'00s","'10s","'20s" ))+
  scale_fill_brewer(palette = "Set2", name = "Pre-40s & Unknown")+
  theme(legend.position = c(.85,.77))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents of Onshore Gas Transmission Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")

```

## Unknown Mileage | A New Issue

```{r unknown phmsa?, include = TRUE}
gt.unknown.group <- gt.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,                 
                                          na = c("NA","UNKNOWN","N/A")),    
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1930),           
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",   
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR !=2022)%>%
  left_join(gtMileKnown,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T)%>%                                                        
  #filter(!is.na(value))%>%                                                  
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),                                                         
            miles = mean(value),                                             
            s.ipm = n/(miles/1000))%>%    
  add_row(INSTALLATION_DECADE = 1920, s.ipm = 0.0)%>%
  mutate(version = "Grouped")
  
  
gt.unknown.split <- gt.inc %>%
  #drop_na(INSTALLATION_YEAR)%>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,na = c("NA","UNKNOWN")),             
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1930, 1930, INSTALLATION_DECADE),
         INSTALLATION_DECADE = replace_na(INSTALLATION_DECADE, 1920))%>%
  filter(SIGNIFICANT == "YES",   
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR !=2022)%>%
  left_join(gtMileKnown,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T)  %>%                                                   
  #filter(!is.na(value))%>%                                                  
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>% 
  mutate(version = "Separate")

rbind(gt.unknown.split, gt.unknown.group)%>%   
  mutate(s.ipm = if_else(INSTALLATION_DECADE == 2020, s.ipm*5, s.ipm))%>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm, fill = version, group = version))+ 
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.5)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1915,2025),
                     breaks = seq(1920, 2020,10),
                     labels = c("Unknown", "Pre-40s", 
                                paste0("'",seq(40,90,10),"s"),
                                "'00s","'10s","'20s" ))+
  scale_fill_brewer(palette = "Set2", name = "Pre-40s & Unknown")+
  theme(legend.position = c(.85,.77))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents of Onshore Gas Transmission Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")
```

## Unknown Mileage | Why?

- In the original, I was separately averaging incidents per decade-year mileage  
    - The revision is summing incidents and summing mileage to get a rate   
    - Issues arise when you divide by 0   
- 0 Appears in 2010 and 2011 (not in the original plot's time frame)  
    -  In these years, Pre-40s and Unknown Mileage was ALL Unknown  
    -  Despite this, a handful of incidents in these years had pre-40s install year dates   
    -  The old method to get averages would result in infinite values and an infinite average  


# Hazardous Liquids

## Considerations | System Parts

```{r include = TRUE}
table(hl.inc$SYSTEM_PART_INVOLVED)
```

## Base Plot 

```{r hl base plot, include = TRUE}
hl.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1910),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1910, 1910, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR < 2020)%>%
  left_join(hlMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(s.ipm = if_else(INSTALLATION_DECADE == 2020, s.ipm*5, s.ipm))%>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm))+ 
  geom_bar(stat = "identity", fill = midBlue)+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.7)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1905,2015),
                     breaks = seq(1910, 2010,20),
                     labels = c("Pre-20s\n& Unkown",seq(1930,2010,20)))+
  scale_y_continuous(expand = expansion(mult = c(0,.1), add=0))+
  scale_fill_brewer(palette = "Paired", direction = -1)+
  theme(legend.position = c(.7,.7))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents of Onshore Hazardous Liquid Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")
```


## Adding Parts

```{r hl add parts, include = TRUE}
hl.base <- hl.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1910),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1910, 1910, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" ,
         IYEAR < 2020)%>%
  left_join(hlMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "Base")

hl.offshore <- hl.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1910),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1910, 1910, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         SYSTEM_PART_INVOLVED == "ONSHORE PIPELINE, INCLUDING VALVE SITES" |
          SYSTEM_PART_INVOLVED == "OFFSHORE PIPELINE, INCLUDING RISER AND RISER BEND",
         IYEAR < 2020)%>%
  left_join(hlMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ Offshore")

hl.piping <- hl.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1910),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1910, 1910, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         grepl("PIPING", SYSTEM_PART_INVOLVED),
         IYEAR < 2020)%>%
  left_join(hlMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ 'Piping'")

hl.all <- hl.inc %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR,
                                          na = c("NA","UNKNOWN","N/A")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1910),
         INSTALLATION_DECADE = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10 ,
         INSTALLATION_DECADE = if_else(INSTALLATION_DECADE <= 1910, 1910, INSTALLATION_DECADE))%>%
  filter(SIGNIFICANT == "YES",
         IYEAR < 2020)%>%
  left_join(hlMileNorm,                   
            by = c("INSTALLATION_DECADE" = "name", "IYEAR" = "Calendar.Year"),             
            keep = T) %>%
  group_by(INSTALLATION_DECADE)%>%
  summarise(n = n(),
            miles = mean(value),
            s.ipm = n/(miles/1000))%>%
  mutate(version = "+ Everything")

rbind(hl.base, hl.offshore,hl.piping ,hl.all)%>%
  mutate(version = factor(version, levels = c("Base", "+ Offshore","+ 'Piping'", "+ Everything"))) %>%
  ggplot(aes(x = INSTALLATION_DECADE, y = s.ipm, fill = version, group = version))+ 
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = round(s.ipm,1), family = pstFont), 
            vjust = 1.5, colour = "white", position = position_dodge(width = 9.3), size = 2.3)+
  theme_pst(baseSize = 12)+
  scale_x_continuous(limits = c(1905,2015),
                     breaks = seq(1910, 2010,20),
                     labels = c("Pre-20s\n& Unkown",seq(1930,2010,20)))+
  scale_y_continuous(expand = expansion(mult = c(0,.1), add=0))+
  scale_fill_manual(name = "Parts", values = c(darkBlue, midBlue, bluGreen, liteGreen))+
  theme(legend.position = c(.7,.7))+
  labs(title = "Pipeline Incidents Per 1,000 Miles by Installation Decade",
       subtitle = "Significant Incidents of Onshore Hazardous Liquid Lines from the 1940s to Present",
       caption = capt,
       tag = "PST 2022",
       y = "Incidents per 1,000 Miles",
       x = "Installation Decade")
```

