---
output: html_document
editor_options: 
  chunk_output_type: console
---
 ---
title: "Untitled Analysis"
subtitle: "Specifics (Timeline)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout: default
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{img/PST_logo.png}\LARGE\\}
  - \posttitle{\end{center}}
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE, echo = FALSE}
rm(list = ls())
```

```{r setup, include=FALSE}
library(tufte)
library(incays)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(lubridate)
library(sp)
library(sf)
library(rgeos)
library(tufte)
library(rgdal)
library(ggspatial)
library(mapboxapi)

knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
options(knitr.table.format = "latex")
capt = "Source: PHMSA Incident and Mileage Data (2010-2021)"
datadir = "https://raw.githubusercontent.com/jmceager/pst/main/phmsa-clean/data/"
tag = paste("PST", lubridate::year(Sys.Date()))
##
moState = "MN"
moStateL = state.name[match(moState,state.abb)]
```

```{r theme_pst}
library(showtext)
#> Loading required package: sysfonts
#> Loading required package: showtextdb
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()
theme_pst <- function(baseSize=8) {
  (theme(
                       axis.line.x = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.line.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.ticks.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       text = element_text(family = pstFont),
                       axis.text = element_text(colour = "#8C9394",
                                                size = baseSize * .8),
                       axis.title = element_text(colour ="black"),
                       panel.grid.minor.y = element_line(linetype = "dotted", 
                                                         colour = "#C4C8C6"),
                       panel.grid.major.y = element_line(colour = "#394C56", 
                                                         linetype = "dotted"),
                       panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(),
                       panel.background = element_blank(),
                       panel.border = element_blank(),
                       strip.background = element_rect(linetype = 0),
                       strip.text = element_blank(),
                       strip.text.x = element_blank(),
                       strip.text.y = element_blank(),
                       legend.text = element_text(colour = "black",
                                                  size = baseSize * .8),
                       legend.background = element_rect(fill = NULL, 
                                                        color = "#182125"),
                       legend.title = element_text( face = "bold", 
                                                    colour = "black",
                                                    size = baseSize),
                       legend.position = "right",
                       legend.key = element_blank(),
                       #legend.background = element_blank(),
                       plot.tag.position = "topright",
                       plot.tag = element_text(size = baseSize *.66,
                                               colour = "#8C9394"),
                       plot.background = element_rect(fill = "#ffffff"),
                       plot.title = element_text(face = "bold", 
                                                 colour = "black",
                                                 size = baseSize*1.2),
                       plot.subtitle = element_text(colour = "#8C9394",
                                                    size = baseSize)
                     )
  )
}
```

```{r nice colors}
#brewer pal
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]

#pst 
darkBlue <- "#003E59"
midBlue <- "#00546F"
bluGreen <- "#459197"
liteGreen <- "#61A893"
```

```{r loading incident data}
#all incident data + op mileage
df <- read_csv(paste0(datadir, "clean/all_inc.csv")) %>%
  filter(STATE == moState)

adj.states <- c("SD","WI", "ND", "IA") #insert adjacent states here 

df.adj <- read_csv(paste0(datadir, "clean/all_inc.csv"))%>%
  filter(STATE %in% adj.states)%>%
  mutate(MSYS = str_sub(SYS, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

df.all <- read_csv(paste0(datadir, "clean/all_inc.csv"))
```


From the rbind codebloack below, MN is most similar to the US average in terms of age of pipeline involved in incidents. The most common age of failed pipelines are unknown in both cases; while of known pipelines, the average age is in the mid-1980s and the median in the 1980s. The average yearly incidents for MN and the US is also similar: 136 to 146. 
```{r}
read_csv(paste0(datadir, "clean/all_inc.csv")) %>%
  group_by(STATE)%>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR, na = "UNKNOWN"),
         mean = mean(INSTALLATION_YEAR, na.rm = T),
         median = median(INSTALLATION_YEAR, na.rm = T))%>%
  ungroup()%>%
  mutate(INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1900))%>%
  group_by(STATE) %>%
  summarise(mean = max(mean),
            mode = modeFind(INSTALLATION_YEAR),
            median = max(median),
            n = n()) %>% 
  distinct(STATE, .keep_all = T) %>%
 # filter(STATE == "MN")%>%
  rbind(read_csv(paste0(datadir, "clean/all_inc.csv")) %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR, na = "UNKNOWN"),
         mean = mean(INSTALLATION_YEAR, na.rm = T),
         median = median(INSTALLATION_YEAR, na.rm = T),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1900))%>%
  summarise(mode = modeFind(INSTALLATION_YEAR),
            mean = max(mean),
            median = max(median),
            n = n()/52,
            STATE = "U.S.A.") ) %>% view()



read_csv(paste0(datadir, "clean/all_inc.csv")) %>%
  mutate(INSTALLATION_YEAR = parse_number(INSTALLATION_YEAR, na = c("UNKNOWN", "NA")),
         INSTALLATION_YEAR = replace_na(INSTALLATION_YEAR, 1900),
         IDEC = INSTALLATION_YEAR - INSTALLATION_YEAR %% 10) %>%
  group_by(IDEC)%>%
  summarize(n = n())
```


```{r mileage data}
miles <- read_csv(paste0(datadir, "clean/sys_miles.csv")) %>%
  group_by(STATE, IYEAR) %>%
  summarise(miles = sum(mileage))

stateMiles <- filter(miles, STATE == moState) %>% select(miles, IYEAR)

milesFull <- read_csv(paste0(datadir, "clean/sys_miles.csv"))

adjSM <- read_csv(paste0(datadir,"clean/sys_miles.csv"))%>%
  filter(IYEAR >=2010,
         STATE == moState | STATE %in% adj.states)%>%
  mutate(STATE = if_else(STATE != moState, "Adjacent (avg.)", STATE))%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  mutate(mileage = if_else(STATE != moState, mileage/length(adj.states), mileage))

allSM <- read_csv(paste0(datadir,"clean/sys_miles.csv")) %>% 
  filter(IYEAR >=2010, STATE != moState)%>%
  mutate(STATE = "Rest of U.S. (avg.)")%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  mutate(mileage = if_else(STATE != moState, mileage/length(unique(df.all$STATE)), mileage)) %>%
  rbind(adjSM)
```

# Pipeline Incident Analysis  

\newthought{This makes it caps} or something to start new sections. Links might not show up properly so use this formatting: \href{https://www.phmsa.dot.gov/data-and-statistics/pipeline/pipeline-incident-flagged-files}{\textcolor{Cerulean}{\underline{PHMSA's website}}}. 

## Overview  
```{r quick facts}
sig <- df %>%
  count(SIGNIFICANT)%>%
  filter(SIGNIFICANT == "YES")

bad <- df %>%
  count(INJURY_IND, FATALITY_IND) %>%
  filter(INJURY_IND != "NO" | FATALITY_IND != "NO") %>%
  summarise(n = sum(n))

fire <- df %>%
  count(EXPLODE_IND, IGNITE_IND) %>%
  filter(EXPLODE_IND != "NO" | IGNITE_IND != "NO") %>%
  summarise(n = sum(n))

death <- df %>%
  mutate(FATAL = replace_na(FATAL,0)) %>%
  summarise(FATAL = sum(FATAL))

injure <- df %>%
  mutate(INJURE = replace_na(INJURE,0)) %>%
  summarise(INJURE = sum(INJURE))

evac <- df %>%
  mutate(NUM_PUB_EVACUATED = replace_na(NUM_PUB_EVACUATED,0)) %>%
  summarise(NUM_PUB_EVACUATED = sum(NUM_PUB_EVACUATED))

rel <- df %>%
  mutate(UNINTENTIONAL_RELEASE = replace_na(UNINTENTIONAL_RELEASE,0),
         INTENTIONAL_RELEASE = replace_na(INTENTIONAL_RELEASE,0),
         TOTAL_RELEASE = UNINTENTIONAL_RELEASE + INTENTIONAL_RELEASE) %>%
  group_by(SYS)%>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

gasRel <- filter(rel, !grepl("HL", SYS)) %>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

liqRel <- filter(rel, grepl("HL", SYS))

modeFind <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

freqOp <- modeFind(df$NAME)

quickFacts <- tibble(
  #names
  var = c("totInc","liqInc","trnInc","dstInc", 
          "sigInc","humanInd","inj","ftl",
          "fire","evac", "gasRel","liqRel",
          "modeOp"),
  #values
  val = c(nrow(df), nrow(filter(df,grepl("HL",SYS))),
          nrow(filter(df,grepl("GT",SYS))), nrow(filter(df,grepl("GD",SYS))),
          sig[[2]], bad[[1]], injure[[1]], death[[1]],
          fire[[1]], comma(evac[[1]]),comma(gasRel[[1]]), comma(liqRel[[2]]),
          freqOp)
  )
```
**Quick Facts**  

From 2010 to today, XXXXXXX has had: 
  
*  `r nrow(df)` total incidents including:  
  +  `r nrow(filter(df,grepl("HL",SYSTEM_TYPE)))` Hazardous Liquid incidents
  +  `r nrow(filter(df,grepl("GT",SYSTEM_TYPE)))` Gas Transmission incidents  
  +  `r nrow(filter(df,grepl("GD",SYSTEM_TYPE)))` Gas Distribution incidents  
*  `r sig[[2]]` significant incidents
*  `r bad[[1]]` incidents ending in injury and/or fatality
*  `r fire[[1]]` incidents resulting in fire and/or explosion
*  `r comma(evac[[1]])` total members of the public evacuated 
*  `r comma(rel[[1]])` Thousand Cubic Feet (mscf) of gas released as a result of incidents

```{marginfigure include = T, echo = T}
Place margin text like this 
```

```{r}
milesStates <- milesFull %>%
  filter(IYEAR >= 2010)%>%
  group_by(STATE, IYEAR)%>%
  summarise(miles = sum(mileage))

incStates <- df.all %>%
  filter(IYEAR <=2021)%>%
  count(STATE, IYEAR)%>%
  complete(STATE, IYEAR, fill = list(n = 0)) %>%
  left_join(milesStates, by = c("STATE", "IYEAR"))%>%
  mutate(ipm = n / (miles / 1000))%>%
  group_by(STATE)%>%
  summarise(ipm = mean(ipm))%>%
  slice_max(order_by = ipm, n = 10)

axcol <- rev(if_else(incStates$STATE == moState,liteGreen,"gray35"))

incStates %>%
  mutate(interest = if_else(STATE == moState,T,F))%>%
  ggplot(aes(x = reorder(STATE, ipm), y = ipm, color = interest))+
  geom_point(size = 3)+
  theme_pst(baseSize = 12)+
  guides(color = "none")+
  scale_color_manual(values = c("gray35", liteGreen))+
  scale_x_discrete(labels=c("ND"=expression(bold(ND)), parse=TRUE))+
  scale_y_continuous(limits = c(0.25,1.3),
                     breaks = seq(.5,1.5,.5))+
  theme(axis.text.y = element_text(colour = axcol),
        panel.grid.minor.x = element_line(linetype = "dotted", 
                                          colour = "#C4C8C6"),
        panel.grid.major.x = element_line(colour = "#394C56", 
                                          linetype = "dotted"),
        panel.grid.major.y = element_line(colour = "gray85", 
                                          linetype = "dotted"),
        panel.grid.minor.y = element_blank())+
  labs(x = NULL, y = "Mean Incident Rate (per 1,000 Miles)",
       title = "Mean Incident Rate By State",
       subtitle = "Highest 10 States: All Systems from 2010-2021",
       tag = tag, 
       caption = capt)+
  coord_flip()

ggsave("./MN/IPMComp.png", width = wdt, height = hgt, units = "px")
```


## Incident Breakdown

```{r inc rate}
df %>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(df.adj, IYEAR), n = n/6, STATE = "Adjacent (avg.)"),
    mutate(count(df.all, IYEAR), n = n/length(unique(df.all$STATE)), STATE = "Rest of U.S. (avg.)")
    ) %>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/(mileage / 1000))%>%
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  geom_point(aes(size = n), alpha = .8)+
  geom_line()+
  theme_pst()+
  scale_color_manual(values = c(liteGreen, bluGreen, darkBlue),
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "Rest of U.S. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "Rest of U.S. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(0,.6),
                     breaks = seq(0,.8,.2))+ 
  theme(legend.position = c(.85, .9),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Pipeline Incident Rate"),
       subtitle = "All Systems, All Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Incident Count")

wdt = 2560
hgt = wdt/1.62
ggsave("./MN/AllIncMile.png", width = wdt, height = hgt, units = "px")

```


```{r sig per mile}
df %>%
  filter(SIGNIFICANT == "YES")%>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(filter(df.adj, SIGNIFICANT == "YES"), IYEAR), n = n/6, STATE = "Adjacent (avg.)"),
    mutate(count(filter(df.all, SIGNIFICANT == "YES"), IYEAR), n = n/length(unique(df.all$STATE)), STATE = "Rest of U.S. (avg.)")
    )%>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/(mileage/1000) )%>% 
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  geom_point(aes(size = n), alpha = .8)+
  geom_line()+
  theme_pst()+
  scale_color_manual(values = c(liteGreen, bluGreen, darkBlue),
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "Rest of U.S. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "Rest of U.S. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(0, .45),breaks = seq(0,.45,.15))+ 
  theme(legend.position = c(.8, .8),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Significant Incident Rate"),
       subtitle = "All Systems, Significant Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Significant Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Singificant\nIncident Count")

wdt = 2560
hgt = wdt/1.62
ggsave("./MN/SigIncMile.png", width = wdt, height = hgt, units = "px")
```


## Operator Analysis 

```{r op table}
milesND <- milesFull %>%
  filter(OPERATOR_ID %in% unique(df$OPERATOR_ID),
         STATE == moState) %>%
  group_by(OPERATOR_ID)%>%
  summarise(mileage = mean(mileage))


#note: figure out how to more accurately rep ipm 
df.t <- df %>%
  left_join(milesND, by = "OPERATOR_ID")%>%
  mutate(SIGNIFICANT = if_else(SIGNIFICANT == "YES", 1,0),
         EXPLODE_IND = if_else(EXPLODE_IND == "YES", 1,0),
         IGNITE_IND = if_else(IGNITE_IND == "YES", 1,0)) %>%
  group_by(pri.name, SYS)%>%
  summarise(miles = mean(mileage.y),
            n = n(),
            sig = sum(SIGNIFICANT)/n,
            ipm = n/(miles/1000),
            trend = "",
            fatal = sum(FATAL),
            injure = sum(INJURE),
            evac = sum(NUM_PUB_EVACUATED, na.rm = T),
            cost = sum(TOTAL_COST_CURRENT)/1000,
            cpi = cost/ n ,
            rel = sum(TOTAL_RELEASE),
            rpi = rel/n )%>%
  ungroup()

t.gd <- df.t %>%
  filter(SYS == "GD") %>%
  mutate(pri.name = if_else(grepl("CENTERPOINT", pri.name), "Centerpoint (GD)", pri.name),
         pri.name = fct_lump(f=pri.name, 4, w=n, other_level = "Others (GD, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

t.gt <- df.t %>%
  filter(SYS == "GT") %>%
  mutate(pri.name = if_else(grepl("CENTERPOINT", pri.name), "Centerpoint (GT)", pri.name),
         pri.name = fct_lump(f=pri.name, 4, w=n, other_level = "Others (GT, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

t.hl <- df.t %>%
  filter(SYS == "HL") %>%
  mutate(pri.name = fct_lump(f=pri.name, 4, w=n, other_level = "Others (HL, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T)%>%
  arrange(desc(n))

df.k <- rbind(t.hl, t.gd, t.gt) 

bigNames <- as.character(c(as.character(t.hl$pri.name), as.character(t.gd$pri.name), as.character(t.gt$pri.name)))
bigNames <- bigNames[!grepl("Others",bigNames)]

namesMatch <- tibble(pri.name = bigNames, SYS = c(rep("HL",4),
                                                  rep("GD",4),
                                                  rep("GT",5)))

df.trend <- df %>%
  filter(IYEAR <2022,
         SYS != "UNGS",
         SYS != "GG")%>%
  left_join(namesMatch, by = c("pri.name", "SYS"), keep = T)%>%
  mutate(pri.name.y = if_else(pri.name.x == "CENTERPOINT ENERGY RESOURCES CORP., DBA CENTERPOINT ENERGY MINNESOTA GAS",
                              if_else(SYS.x == "GT", "Centerpoint (GT)", "Centerpoint (GD)"),
                              pri.name.y),
         pri.name.y = if_else(is.na(pri.name.y),
                              if_else(SYS.x == "GT", "Others (GT, avg.)",
                                      if_else(SYS.x == "GD", "Others (GD, avg.)", 
                                              if_else(SYS.x == "HL", "Others (HL, avg.)",
                                                      "N"))),
                              pri.name.y),
         SYS.y = coalesce(SYS.y, SYS.x))%>%
  rename(pri.name = pri.name.y,
         SYS = SYS.y)%>%
  count(pri.name, SYS, IYEAR)%>%
  complete(pri.name, SYS, IYEAR, fill = list(n = 0))%>%
  left_join(namesMatch %>% mutate(pri.name = if_else(grepl("CENTERPOINT",pri.name),
                                                     if_else(SYS == "GT", "Centerpoint (GT)", "Centerpoint (GD)"),
                                                     pri.name))
            , by = c("pri.name", "SYS"), keep = T) %>%
  filter(!is.na(pri.name.y) | grepl("Others",pri.name.x))%>%
  mutate(drop = if_else(!str_detect(pri.name.x,SYS.x) & 
                          (grepl("Others",pri.name.x)| grepl("Centerpoint", pri.name.x)), T, F))%>%
  filter(!drop)%>%
  rename(pri.name = pri.name.y)%>%
  group_by(pri.name)%>%
  mutate(tot = sum(n),
         n = (n/tot)*100,
         pri.name = if_else(is.na(pri.name), pri.name.x, pri.name))%>%
  ungroup()%>%
  mutate(pri.name = factor(pri.name, levels = levels(df.k$pri.name)))

df.lines <- split(df.trend$n, df.trend$pri.name)


df.k %>%
  select(!c(fatal, injure,evac)) %>%
  mutate(trend = "")%>%
  mutate(n = round(n,0),
         sig = paste0(round(sig*100, 0), "%"))%>%
  mutate_at(vars(cost:rpi), round, 0)%>%
  kbl(col.names = c("Parent Operator", "Total", "Percent Significant",
                    "per 1K Miles", "Trend","Total","per Incident", "Total","per Incident"),
      format.args = list(big.mark = ","), digits = 2, booktabs = TRUE)%>%
  kable_classic_2(lightable_options = c("striped","hover"))%>%
  add_header_above(c(" "=1, "Incidents" = 4, "Cost\n(K, USD)" =2, "Releases" = 2)) %>%
  column_spec(5, image = spec_plot(df.lines, same_lim = T, 
                                   min = list(col = orange, pch = 20, cex = .2),
                                   max = list(col = green, pch = 20, cex = .2)))%>%
  column_spec(4, color=spec_color(df.k$ipm, option = "B", end = .7), bold = T)%>%
  column_spec(7, color=spec_color(df.k$cpi, option = "B", end = .7), bold = T)%>%
  column_spec(9, color=spec_color(df.k$rpi, option = "B", end = .7), bold = T) %>%
  group_rows(start_row = 1, end_row = 5, group_label = "Hazardous Liquids")%>%
  group_rows(start_row = 6, end_row = 10, group_label = "Gas Distribution")%>%
  group_rows(start_row = 11, end_row = 15, group_label = "Gas Transmission") %>%
  cat(., file = "./MN/table2.html")
```


## Spatial Analysis  
```{r load state shapefile}
projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

stateShape <- read_sf("gis/us_states.shp")

stateShape <- stateShape[stateShape$NAME == moStateL,]

#mb_access_token("accesstokenhere", install = T)
```

```{r map, include = TRUE, fig.fullsize = TRUE}
df.sp <- st_as_sf(df, 
                  coords = c("LOCATION_LONGITUDE","LOCATION_LATITUDE"),
                  crs = projcrs) 

df.state <- st_join(df.sp, st_transform(stateShape, crs = projcrs), join = st_within) %>%
  filter(NAME.y == moStateL) %>%
  st_transform(crs = 3857)

bbox <- st_bbox(df.state, crs = 3857)

base_tiles <- get_static_tiles(location = stateShape,
                               zoom = 5,
                               style_id = "cl1pbimv7000114ouzmsb6obk",
                               username = "jimeagah",
                               buffer_dist = 1500000,
                               scaling_factor = "1x",
                               crop = T)

place_tiles <- get_static_tiles(location = stateShape,
                                zoom = 5,
                                style_id = "cl6zaoczy004n14jqcivrkvb5",
                                username = "jimeagah",
                                buffer_dist = 1000,
                                scaling_factor = "2x",
                                crop = F)



##ggspatial
df.state%>%
  ggplot()+
  layer_spatial(base_tiles)+
  geom_sf(data = stateShape, fill = NA, color = "gray50")+
  layer_spatial(place_tiles)+
  geom_sf(aes(color = SYS, size = TOTAL_COST_CURRENT), alpha = .6)+
  scale_size(name = "Cost of Damage\n(Mil. USD)",
             breaks = seq(0,40000000, 10000000),
             labels = seq(0,40,10))+
  scale_color_brewer(palette = "Set2", name = "System",
                     labels = c("Gas Distribution","Gas Transmission", "Hazardous Liquids"))+
  coord_sf(crs = 3857,
           xlim = c(bbox[[1]] - 500000  , bbox[[3]] + 500000),
           ylim = c(bbox[[2]] , bbox[[4]]+50000 ),
           expand = TRUE)+
  labs(title = paste0("Incident Locations in ", moStateL),
       subtitle = "All Systems (2010 - 2021)",
       x = NULL,
       y = NULL,
       caption = capt,
       tag = tag)+
  theme_void()+
  theme(legend.position = c(0.1,0.66),
        text = element_text(pstFont),
        plot.margin = unit(c(0,0,0,0), "cm"),
        plot.tag.position = c(.96,0.95),
        plot.tag = element_text(color = "gray 30", size = 8),
        plot.title = element_text(face = "bold", colour = "black", hjust = .015),
        plot.subtitle = element_text(colour = "black", hjust = .015),
        plot.caption = element_text(colour = "gray 30"),
        legend.title = element_text(colour = "gray 10", face = "bold"),
        legend.text = element_text(colour = "gray 10"),
        plot.background = element_rect(fill = "#E8EBEA", colour = NA)) #fill is to match bg

wdt = 2560
hgt = wdt/1.62
ggsave("MN/map.png", width = wdt, height = hgt, units = "px")
``` 

```{r}
df %>% 
  rbind(df.all %>% filter(STATE != "MN") %>% mutate(STATE = "Rest of U.S."))%>%
  group_by(STATE)%>%
  summarise(total = n()/12,
            significant = (sum(SIGNIFICANT == "YES")/total/12)*100,
            serious = sum(SERIOUS == "YES")/total/12*100,
            explode = sum(EXPLODE_IND == "YES")/total/12*100,
            fire = sum(IGNITE_IND == "YES")/total/12*100,
            cost = sum(TOTAL_COST_CURRENT/1000)/12,
            fatal = sum(FATAL)/12,
            injuries = sum(INJURE)/12,
            gd = sum(SYS == "GD")/total,
            gt = sum(SYS == "GT")/total,
            hl = sum(SYS == "HL")/total,
            total = if_else(STATE != "MN", total/50, total))%>%
  distinct(STATE, .keep_all = T)%>%
   kbl(col.names = c("Geography", "Annual Incidents", "Significant",
                    "Serious", "Exploded","Fire","Cost (K)", "Dead","Injured",
                    "GD", "GT", "HL"),
      format.args = list(big.mark = ","), digits = 1, booktabs = TRUE,
      caption = "Minnesota Incident Characteristics Compared to the Rest of the U.S. Average")%>%
  kable_classic_2(lightable_options = c("striped","hover"))%>%
  add_header_above(c(" "=2, "Percent" = 4, "per Year" = 3, "Annual Incidents" = 3))
  cat(., file = "./MN/avg.html")
```

```{r}
df.all %>%
  group_by(STATE)%>%
  summarise(total = n(),
            gd = sum(SYS == "GD")/total,
            injuries = sum(injure)/total,
            MN = if_else(STATE == "MN", "MN","NO"))%>%
  ggplot(aes(x = gd, y = injuries, color = MN))+
  geom_point()
```

```{r}
df.z <- df.all %>%
  group_by(STATE) %>%
  summarise(total = n()/12,
            significant = sum(SIGNIFICANT == "YES")/total,
            serious = sum(SERIOUS == "YES")/total,
            explode = sum(EXPLODE_IND == "YES")/total,
            fire = sum(IGNITE_IND == "YES")/total,
            cost = sum(TOTAL_COST_CURRENT)/total,
            fatal = sum(FATAL)/total,
            injuries = sum(injure)/total,
            gd = sum(SYS == "GD")/total,
            gt = sum(SYS == "GT")/total,
            hl = sum(SYS == "HL")/total)%>%
  mutate(across(where(is.numeric), scale))%>%
  pivot_longer(!STATE)%>%
  mutate(mn = if_else(STATE == "MN", T,F),
         name = factor(name, levels = c("total","significant","serious",
                                        "fire","explode","injuries","fatal", "cost",
                                        "hl","gd","gt")))

df.z %>%
  ggplot(aes(y = value, x = name, label = STATE))+
  geom_boxplot(outlier.shape = NA, color = midBlue, weight = 8)+
  geom_jitter( data = filter(df.z, STATE != "MN"),
               aes(color = mn), width =.2, alpha = .6)+
  geom_point(data = filter(df.z, STATE == "MN"),
             aes(y = value, x = name, color = mn), size = 2.5)+
  scale_x_discrete(
    labels = rev(c("Incidents","Significant","Serious",
                              "Fire","Explosion", "Injuries", "Fatalities","Cost",
                              "HL","GT","GT")),
                   limits = rev(levels(df.z$name))
                   )+
  scale_color_manual(values = c(darkBlue, liteGreen), 
                     labels = c("Other","Minnesota"),
                     name = "State")+
  theme_pst()+
  theme(panel.grid.minor.x = element_line(linetype = "dotted", 
                                          colour = "#C4C8C6"),
        panel.grid.major.x = element_line(colour = "#394C56", 
                                          linetype = "dotted"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = c(0.1,0.7))+
  labs(title = "State Incident Characteristic Z-Scores",
       subtitle = "All Systems from 2010-2021",
       tag = tag,
       caption = capt,
       x = "Variable",
       y = "Z-score")+
  coord_flip()

wdt = 2560
hgt = wdt/1.62
ggsave("MN/box.png", width = wdt, height = hgt, units = "px")
```

```{r}
df.release <- df %>%
  filter(MSYS == "HL") %>%
  mutate( TOTAL_RELEASE = scale(TOTAL_RELEASE) ) %>%
  rbind(  
    filter(df, MSYS == "Gas") %>% mutate(TOTAL_RELEASE = scale(TOTAL_RELEASE))
    ) %>%
  mutate(SIGSER = if_else(SIGNIFICANT == "YES", 
                          if_else(SERIOUS == "YES", 2,1),
                          0),
         pri.name = if_else(grepl("CENTERPOINT", pri.name), 
                            paste0("Centerpoint (", SYS, ")"),
                            pri.name))

df.release %>%
  ggplot(aes(x = MDY, y = TOTAL_RELEASE, size = TOTAL_COST_CURRENT/1000000, color = as.factor(SIGSER)))+
  geom_point(alpha = .9)+
  geom_label_repel(data = filter(df.release, TOTAL_RELEASE > 3 | SIGSER > 1),
                  aes(x = MDY, y = TOTAL_RELEASE, label = pri.name),
                  size = 4, color = "gray 20",  parse = F,
                  box.padding = .6, min.segment.length = .4, 
                  nudge_y = -0.2, nudge_x = 15, seed = 2020,)+
 # annotate("text", x = ymd("2010-12-30"), y = 8, label = "ok")+
  theme_pst()+
  theme(legend.position = c(.9,.6))+
  scale_color_brewer(palette = "Purples", name = "Significance",
                     labels = c("None", "Significant","Serious"))+
  scale_y_continuous(limits = c(-1,8.5))+
  scale_size(breaks = seq(0,50,10), 
             labels =scales::unit_format(prefix = "$",
                                         suffix = "M") ,
             name = "Damage Cost (2022 USD)")+
  labs(caption = paste0("Release Z-Scores are specific to unit of system measurement: MSCF for Natural Gas and BBLs for Liquids." ,"\n",capt),
       tag = tag,
       x = "Date",
       y = "Release Z-Score",
       title = "Minnesota Pipeline Incident Detail",
       subtitle = "Total Release Z-Scores, Damage Cost, and Significance (2010-2021)")

ggsave("MN/op.png", width = wdt, height = hgt, units = "px")
```




