---
title: "Untitled Analysis"
subtitle: "Specifics (Timeline)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout: default
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{img/PST_logo.png}\LARGE\\}
  - \posttitle{\end{center}}
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE, echo = FALSE}
rm(list = ls())
```

```{r setup, include=FALSE}
#load tufte, basis for package
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

#load libs
library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(sp)
library(rgeos)
library(rgdal)
library(ggspatial)
library(mapboxapi)
library(showtext)

#setting state
moState = "NC"
moStateL = state.name[match(moState,state.abb)]

# setting options 
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
options(knitr.table.format = "html")

#setting plot font 
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()

#setting plot items
capt = "Source: PHMSA Incident and Mileage Data (2010-2022)"
datadir = "https://raw.githubusercontent.com/jmceager/pst/main/phmsa-clean/data/"
tag = paste("PST", lubridate::year(Sys.Date()))
bs = 12
exportdir = paste0("./", moState, "/" )
w = 8
h = w/1.62
u = "in"
```

```{r theme_pst}

#### map theme ####
theme_pst_map <- function(baseSize=bs) {
  (theme_void() + 
     theme(
       legend.direction = "horizontal",
       legend.position = "bottom",
       plot.tag.position = c(.97,.97),
       text = element_text(family = pstFont),
       legend.text = element_text(colour = "black",
                                   size = baseSize * .8),
       legend.background = element_rect(fill = NA, 
                                        color = NA),
       legend.title = element_text( face = "bold", 
                                    colour = "black",
                                    size = baseSize),
       legend.key = element_blank(),
       plot.background = element_rect(fill = "#ffffff", colour = NA),
       plot.margin = unit(c(0,0,0,0), "cm"),
       plot.tag = element_text(size = baseSize *.66,
                               colour = "#8C9394"),
       plot.title = element_text(face = "bold", 
                                 colour = "black",
                                 size = baseSize*1.2),
       plot.subtitle = element_text(colour = "#8C9394",
                                    size = baseSize),
       plot.caption = element_text(size = baseSize *.8,
                                   lineheight = .33)
       )

                    
  )
}

# theme clean adjustments to apply to ggplot
theme_pst <- function(font = pstFont, fontSize = bs) {
  (theme(
    axis.line.x = element_line(
      colour = "#B3C7D0",
      linewidth = 0.5,
      linetype = "solid"
    ),
    axis.line.y = element_line(
      colour = "#B3C7D0",
      linewidth = 0.5,
      linetype = "solid"
    ),
    axis.ticks.y = element_line(
      colour = "#B3C7D0",
      linewidth = 0.5,
      linetype = "solid"
    ),
    text = element_text(family = font,
                        size = fontSize),
    axis.text = element_text(colour = "#81A2B1",
                             size = fontSize),
    axis.title = element_text(colour ="#C4C8C6",
                              size = fontSize),
    panel.grid.minor.y = element_line(linetype = "dotted", 
                                      colour = "#557A8B",
                                      linewidth = .25),
    panel.grid.major.y = element_line(colour = "#81A2B1", 
                                      linetype = "dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank(),
    strip.background = element_rect(linetype = 0),
    strip.text = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_blank(),
    legend.text = element_text(colour = "#C4C8C6",
                               size = fontSize * .8),
    legend.background = element_rect(fill = "#233239", color = "#B3C7D0"),
    legend.title = element_text( face = "bold", 
                                 colour = "#C4C8C6",
                                 size = fontSize,
                                 hjust = 0),
    legend.position = "right",
    legend.key = element_blank(),
    legend.margin = margin(4,12,4,6, "pt"),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "#233239"),
    plot.title = element_text(face = "bold", 
                              colour = "#C4C8C6",
                              size = fontSize * 1.2,
                              hjust = 0),
    plot.subtitle = element_text(colour = "#C4C8C6",
                                 size = fontSize,
                                 hjust = 0)
  )
  )
}


theme_pst_lite <- function(font = pstFont, fontSize = bs) {
  (theme(
    axis.line.x = element_line(
      colour = "#233239",
      linewidth = 0.5,
      linetype = "solid"
    ),
    axis.ticks.x = element_line(
      colour = "#233239",
      linewidth = 0.5,
      linetype = "solid"
    ),
    axis.line.y = element_line(
      colour = "#233239",
      linewidth = 0.5,
      linetype = "solid"
    ),
    axis.ticks.y = element_line(
      colour = "#233239",
      linewidth = 0.5,
      linetype = "solid"
    ),
    text = element_text(family = font,
                        size = fontSize),
    axis.text = element_text(colour = "#233239",
                             size = fontSize),
    axis.title = element_text(colour ="#001D29",
                              size = fontSize),
    panel.grid.minor.y = element_line(linetype = "dotted", 
                                      colour = "#4D6B80",
                                      linewidth = .25),
    panel.grid.major.y = element_line(colour = "#233239", 
                                      linetype = "dotted",
                                      linewidth = .5),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank(),
    strip.background = element_rect(linetype = 0),
    strip.text = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_blank(),
    legend.text = element_text(colour = "#001D29",
                               size = fontSize * .8),
    legend.background = element_rect(fill = "#FFFFFF", color = "#001D29"),
    legend.title = element_text( face = "bold", 
                                 colour = "#001D29",
                                 size = fontSize,
                                 hjust = 0),
    legend.position = "right",
    legend.key = element_blank(),
    legend.margin = margin(4,12,4,6, "pt"),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "#FFFFFF"),
    plot.title = element_text(face = "bold", 
                              colour = "#001D29",
                              size = fontSize * 1.2,
                              hjust = 0,
                              vjust = 1),
    plot.subtitle = element_text(colour = "#001D29",
                                 size = fontSize,
                                 hjust = 0),
    plot.tag = element_text(size = fontSize *.66, 
                            colour = "#4D6B80",
                            hjust = 1,
                            vjust = 0),
    plot.tag.position = c(.99,.99)
  )
  )
}

```

```{r nice colors}
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]

#brewer pal
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]

#pst 
darkBlue <- "#003E59"
midBlue <- "#00546F"
bluGreen <- "#459197"
liteGreen <- "#61A893"

#pst contrast 
#1
honey <- "#FFAD05"
lilac <- "#C297B8"
crimson <- "#92140C"

# 2
hunterGrn <- "#315C2B"
maize <- "#E7BB41"

#pals 
pst1 <- c(darkBlue, liteGreen, lilac, bluGreen, honey,  crimson)
pst2 <- c(darkBlue, liteGreen, hunterGrn, bluGreen, maize, midBlue)
```

```{r loading incident data}
#all incident data + op mileage
df <- read_csv(paste0(datadir, "clean/all_inc.csv")) %>%
  filter(STATE == moState,
         IYEAR < 2023)

adj.states <- c("VA", "GA", "SC", "TN") #insert adjacent states here 

df.adj <- read_csv(paste0(datadir, "clean/all_inc.csv"))%>%
  filter(STATE %in% adj.states)%>%
  mutate(MSYS = str_sub(SYS, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

df.all <- read_csv(paste0(datadir, "clean/all_inc.csv"))
```

```{r mileage data}
#all miles
miles <- read_csv(paste0(datadir, "clean/sys_miles.csv")) %>%
  group_by(STATE, IYEAR) %>%
  summarise(miles = sum(mileage))

#summaries of miles
stateMiles <- filter(miles, STATE == moState) %>% select(miles)
stateMilesYear <- filter(miles, STATE == moState) %>% select(miles, IYEAR)

#full miles data
milesFull <- read_csv(paste0(datadir, "clean/sys_miles.csv"))

#adjacent state miles
adjSM <- milesFull%>%
  filter(IYEAR >=2010,
         STATE == moState | STATE %in% adj.states)%>%
  mutate(STATE = if_else(STATE != moState, "Adjacent (avg.)", STATE))%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  mutate(mileage = if_else(STATE != moState, mileage/length(adj.states), mileage))

#other state miles
allSM <- milesFull %>% 
  filter(IYEAR >=2010)%>%
  mutate(STATE = "U.S.A. (avg.)")%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  rbind(adjSM)
```


```{r 2021 to 22 incidents}
#dfFull <- read_csv(paste0(datadir, "clean/all_inc.csv")) 

dfFull%>% 
  filter(IYEAR > 2020,
         IYEAR != 2023) %>%
  count(STATE, IYEAR) %>%
  pivot_wider(names_from = IYEAR,
              values_from = n) %>%
  mutate(change = (`2022`-`2021`),
         per_change = change/`2021`) %>%
  arrange(-change)%>%
  top_n(n = 8, wt = change) %>%
  mutate(sLong = state.name[match(STATE,state.abb)])%>%
  ggplot(aes(x = reorder(sLong, change))) +
  geom_point(aes(size = `2022`, y = change), col = darkBlue)+
  geom_linerange(aes(ymin = 0, ymax = change), col = darkBlue)+
  scale_y_continuous(breaks = seq(0,8,4), name = "Incident Increase",
                     expand = expansion(mult = c(0,0.1)))+
  scale_size(name = "2022 Incidents")+
  theme_pst_lite()+
  labs(x = "State",
       title = "Change in Pipeline Incident Count",
       subtitle = "All Systems, 2021 to 2022", 
       caption = capt, tag = tag)+
  coord_flip()+
  theme(legend.position = c(0.8,.2),
        panel.grid.minor.x = element_line(linetype = "dotted", 
                                      colour = "#4D6B80",
                                      linewidth = .25),
    panel.grid.major.x = element_line(colour = "#233239", 
                                      linetype = "dotted",
                                      linewidth = .5),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

plotName = "ChangeInc.png"
plotExport = paste0(exportdir, plotName)
wdt = 8
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "in")
```


# Pipeline Incident Analysis  

\newthought{This makes it caps} or something to start new sections. Links might not show up properly so use this formatting: \href{https://www.phmsa.dot.gov/data-and-statistics/pipeline/pipeline-incident-flagged-files}{\textcolor{Cerulean}{\underline{PHMSA's website}}}. 

## Overview  
```{r quick facts}
sig <- df %>%
  count(SIGNIFICANT)%>%
  filter(SIGNIFICANT == "YES")

bad <- df %>%
  count(INJURY_IND, FATALITY_IND) %>%
  filter(INJURY_IND != "NO" | FATALITY_IND != "NO") %>%
  summarise(n = sum(n))

fire <- df %>%
  count(EXPLODE_IND, IGNITE_IND) %>%
  filter(EXPLODE_IND != "NO" | IGNITE_IND != "NO") %>%
  summarise(n = sum(n))

death <- df %>%
  mutate(FATAL = replace_na(FATAL,0)) %>%
  summarise(FATAL = sum(FATAL))

injure <- df %>%
  mutate(INJURE = replace_na(INJURE,0)) %>%
  summarise(INJURE = sum(INJURE))

evac <- df %>%
  mutate(NUM_PUB_EVACUATED = replace_na(NUM_PUB_EVACUATED,0)) %>%
  summarise(NUM_PUB_EVACUATED = sum(NUM_PUB_EVACUATED))

rel <- df %>%
  mutate(UNINTENTIONAL_RELEASE = replace_na(UNINTENTIONAL_RELEASE,0),
         INTENTIONAL_RELEASE = replace_na(INTENTIONAL_RELEASE,0),
         TOTAL_RELEASE = UNINTENTIONAL_RELEASE + INTENTIONAL_RELEASE) %>%
  group_by(SYS)%>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

gasRel <- filter(rel, !grepl("HL", SYS)) %>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

hazRel <- filter(rel, grepl("HL", SYS))

modeFind <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

freqOp <- modeFind(df$NAME)

## turn facts into tibble 
quickFacts <- tibble(
  #names
  var = c("totInc","liqInc","trnInc","dstInc", 
          "sigInc","humanInd","inj","ftl",
          "fire","evac", "gasRel","liqRel",
          "modeOp"),
  #values
  val = c(nrow(df), nrow(filter(df,grepl("HL",SYS))),
          nrow(filter(df,grepl("GT",SYS))), nrow(filter(df,grepl("GD",SYS))),
          sig[[2]], bad[[1]], injure[[1]], death[[1]],
          fire[[1]], comma(evac[[1]]),comma(gasRel[[1]]), comma(hazRel[[2]]),
          freqOp)
  )
```
**Quick Facts**  

From 2010 to today, XXXXXXX has had: 
  
*  `r nrow(df)` total incidents including:  
  +  `r nrow(filter(df,grepl("HL",SYSTEM_TYPE)))` Hazardous Liquid incidents
  +  `r nrow(filter(df,grepl("GT",SYSTEM_TYPE)))` Gas Transmission incidents  
  +  `r nrow(filter(df,grepl("GD",SYSTEM_TYPE)))` Gas Distribution incidents  
*  `r sig[[2]]` significant incidents
*  `r bad[[1]]` incidents ending in injury and/or fatality
*  `r fire[[1]]` incidents resulting in fire and/or explosion
*  `r comma(evac[[1]])` total members of the public evacuated 
*  `r comma(rel[[1]])` Thousand Cubic Feet (mscf) of gas released as a result of incidents

```{marginfigure include = T, echo = T}
Place margin text like this 
```

## Incident Breakdown

```{r inc rate}
df %>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(df.adj, IYEAR), STATE = "Adjacent (avg.)"),
    mutate(count(df.all, IYEAR), STATE = "U.S.A. (avg.)")
    ) %>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/(mileage / 1000))%>%
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  #geom_point( alpha = .8)+
  geom_line()+
  theme_pst_lite()+
  scale_color_manual(values = pst1,
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "U.S.A. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "U.S.A. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
   scale_y_continuous(limits = c(0,.22),
                      breaks = seq(0,.2,.05))+ 
  theme(legend.position = c(.2, .8),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Pipeline Incident Rate"),
       subtitle = "All Systems, All Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Incident Count")

plotName = "AllIncMile.png"
plotExport = paste0(exportdir, plotName)
wdt = 2560
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "px")
```

```{r sig per mile}
df %>%
  filter(SIGNIFICANT == "YES")%>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(filter(df.adj, SIGNIFICANT == "YES"), IYEAR), STATE = "Adjacent (avg.)"),
    mutate(count(filter(df.all, SIGNIFICANT == "YES"), IYEAR), STATE = "U.S.A. (avg.)")
    )%>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/(mileage/1000) )%>%
  complete(STATE, IYEAR, fill = list(ir = 0))%>%
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  geom_line()+
  theme_pst_lite()+
  scale_color_manual(values = pst1,
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "U.S.A. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "U.S.A. (avg.)"))+
  scale_y_continuous(limits = c(0, .15),breaks = seq(0,.15,.05),
                     expand = expansion(c(0,0.05)))+ 
    scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  theme(legend.position = c(.2, .8),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Significant Incident Rate"),
       subtitle = "All Systems, Significant Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Significant Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Singificant\nIncident Count")

plotName = "SigIncMile.png"
plotExport = paste0(exportdir, plotName)
wdt = 2560
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "px")
```

## Operator Analysis 

```{r op table}
milesND <- milesFull %>%
  filter(STATE == moState) %>%
  group_by(OPERATOR_ID, IYEAR, SYS)%>%
  summarise(mileage = mean(mileage),
            NAME = NAME,
            pri.name = pri.name,
            pri.id = pri.id)

df.noinc <- milesND %>%
  filter(!OPERATOR_ID %in% df$OPERATOR_ID)%>%
  select(OPERATOR_ID, pri.name, pri.id, SYS, mileage)%>%
  mutate(SIGNIFICANT = 0,
         EXPLODE_IND = 0,
         IGNITE_IND = 0,
         FATAL = 0, 
         INJURE = 0,
         NUM_PUB_EVACUATED = 0,
         TOTAL_COST_CURRENT = 0,
         TOTAL_RELEASE = 0,
         mileage.j = mileage,
         inc = 0) %>%
  select(!mileage)


newNames <- c("PUBLIC SERVICE CO OF NORTH CAROLINA", "PIEDMONT NATURAL GAS CO INC")

#note: figure out how to more accurately rep ipm 
df.t <- df %>%
  left_join(milesND, by = c("OPERATOR_ID", "IYEAR", "SYS"), 
             suffix = c("",".j"), keep = T)%>%
  distinct(NARRATIVE, .keep_all = T)%>%
  mutate(SIGNIFICANT = if_else(SIGNIFICANT == "YES", 1,0),
         EXPLODE_IND = if_else(EXPLODE_IND == "YES", 1,0),
         IGNITE_IND = if_else(IGNITE_IND == "YES", 1,0),
         inc = 1)%>%
    select(OPERATOR_ID, pri.name, pri.id, IYEAR, SYS, SIGNIFICANT, 
           EXPLODE_IND, IGNITE_IND, FATAL, INJURE, NUM_PUB_EVACUATED, 
           TOTAL_COST_CURRENT, TOTAL_RELEASE, mileage.j, inc)%>%
  rbind(df.noinc) %>%
  group_by(pri.id, SYS)%>%
  summarise(pri.name = pri.name,
            miles = mean(mileage.j, na.rm = T),
            n = sum(inc),
            sig = sum(SIGNIFICANT, na.rm = T)/n,
            ipm = n/(miles/1000),
            trend = "",
            fatal = sum(FATAL, na.rm = T),
            injure = sum(INJURE, na.rm = T),
            evac = sum(NUM_PUB_EVACUATED, na.rm = T),
            cost = sum(TOTAL_COST_CURRENT, na.rm = T)/1000,
            cpi = cost/ n ,
            rel = sum(TOTAL_RELEASE, na.rm = T),
            rpi = rel/n )%>%
  ungroup() %>% 
  mutate(across(where(is.numeric) & !c(pri.id), ~ replace_na(.x, 0))) %>%
  distinct(pri.name, SYS,pri.id, .keep_all = T) %>%
  drop_na(pri.name)

lumpD <- 3
nGD <- nrow(df.t[df.t$SYS == "GD", "pri.name"]) - lumpD
t.gd <- df.t %>%
  filter(SYS == "GD") %>% 
  mutate(pri.name = if_else(pri.name %in% newNames, 
                            paste0(pri.name, " (GD)"), pri.name),
         pri.name = fct_lump(f=pri.name, lumpD, w=n, 
                             other_level = paste0(nGD," Others (GD, avg.)")))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

lumpT <- 4
nGT <- nrow(df.t[df.t$SYS == "GT", "pri.name"]) - lumpT
t.gt <- df.t %>%
  filter(SYS == "GT") %>%
  mutate(pri.name = if_else(pri.name %in% newNames, 
                            paste0(pri.name, " (GT)"), pri.name),
         pri.name = fct_lump(f=pri.name, lumpT, w=n, 
                             other_level = paste0(nGT," Others (GT, avg.)")))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

lumpH <- 3
nHL <- nrow(df.t[df.t$SYS == "HL", "pri.name"]) - lumpH
t.hl <- df.t %>%
  filter(SYS == "HL") %>% 
  mutate(pri.name = fct_lump(f=pri.name, lumpH, w=n, 
                             other_level = paste0(nHL," Others (HL, avg.)")))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T)%>%
  arrange(desc(n))

df.k <- rbind(t.hl, t.gd, t.gt) 

bigNames <- as.character(c(as.character(t.hl$pri.name), as.character(t.gd$pri.name), as.character(t.gt$pri.name)))
bigNames <- bigNames[!grepl("Others",bigNames)]

namesMatch <- tibble(pri.name = bigNames, SYS = c(rep("HL",lumpH),
                                                  rep("GD",lumpD),
                                                  rep("GT",lumpT)))

df.trend <- df %>%
  filter(IYEAR <2022,
         SYS != "UNGS",
         SYS != "GG")%>%
  mutate(pri.name = if_else(pri.name %in% newNames, 
                            paste0(pri.name, " (",SYS,")"), pri.name))%>%
  left_join(namesMatch, by = c("pri.name", "SYS"), keep = T)%>%
  mutate(
         # pri.name.y = if_else(pri.name.x == "PIEDMONT NATURAL GAS CO INC",
         #                      if_else(SYS.x == "GT", 
         #                              "PIEDMONT NATURAL GAS CO INC (GT)", 
         #                              "PIEDMONT NATURAL GAS CO INC (GD)"),
         #                      pri.name.x),
         pri.name.y = if_else(is.na(pri.name.y),
                              if_else(SYS.x == "GT", paste0(nGT," Others (GT, avg.)"),
                                      if_else(SYS.x == "GD", 
                                              paste0(nGD," Others (GD, avg.)"), 
                                              if_else(SYS.x == "HL", 
                                                      paste0(nHL," Others (HL, avg.)"),
                                                      "N"))),
                              pri.name.y),
         SYS.y = coalesce(SYS.y, SYS.x))%>%
  rename(pri.name = pri.name.y,
         SYS = SYS.y)%>%
  count(pri.name, SYS, IYEAR)%>%
  complete(pri.name, SYS, IYEAR, fill = list(n = 0))%>%
  left_join(namesMatch %>% mutate(pri.name = if_else(grepl("CENTERPOINT",pri.name),
                                                     if_else(SYS == "GT", "Centerpoint (GT)", "Centerpoint (GD)"),
                                                     pri.name))
            , by = c("pri.name", "SYS"), keep = T) %>%
  filter(!is.na(pri.name.y) | grepl("Others",pri.name.x))%>%
  mutate(drop = if_else(!str_detect(pri.name.x,SYS.x) & 
                          (grepl("Others",pri.name.x)| grepl("Centerpoint", pri.name.x)), T, F))%>%
  filter(!drop)%>%
  rename(pri.name = pri.name.y)%>%
  group_by(pri.name)%>%
  mutate(tot = sum(n),
         n = (n/tot)*100,
         pri.name = if_else(is.na(pri.name), pri.name.x, pri.name))%>%
  ungroup()%>%
  mutate(pri.name = factor(pri.name, levels = levels(df.k$pri.name))) %>%
  complete(pri.name, IYEAR, fill = list(n=0))%>%
  arrange(pri.name)

df.lines <- split(df.trend$n, df.trend$pri.name)

df.lines[order(df.k$pri.name, levels(df.k$pri.name))]

df.k %>%
  select(!c(fatal, injure,evac)) %>%
  mutate(trend = "")%>%
  mutate(n = round(n,1),
         sig = paste0(round(sig*100, 0), "%"))%>%
  mutate_at(vars(cost:rpi), round, 0) %>%
  #table building
  kbl(col.names = c("Parent Operator", "Total", "Percent Significant",
                    "per 1K Miles", "Trend","Total","per Incident", "Total","per Incident"),
      format.args = list(big.mark = ","), digits = 2, booktabs = TRUE) %>%
  kable_classic_2(lightable_options = c("striped","hover"))%>%
  add_header_above(c(" "=1, "Incidents" = 4, "Cost\n(K, USD)" =2, "Releases" = 2)) %>% 
  # color columns relative to values
  column_spec(5, image = spec_plot(df.lines, same_lim = T, 
                                   min = list(col = orange, pch = 20, cex = .2),
                                   max = list(col = green, pch = 20, cex = .2)))%>%
  column_spec(4, color=spec_color(df.k$ipm, option = "B", end = .7), bold = T)%>%
  column_spec(7, color=spec_color(df.k$cpi, option = "B", end = .7), bold = T)%>%
  column_spec(9, color=spec_color(df.k$rpi, option = "B", end = .7), bold = T) %>%
  #group rows according to system
  # TODO: automate start/ends? 
  group_rows(start_row = 1, end_row = lumpH+1, group_label = "Hazardous Liquids")%>%
  group_rows(start_row = lumpH+2, end_row = lumpH+2+lumpD, group_label = "Gas Distribution")%>%
  group_rows(start_row = lumpH+3+lumpD, end_row = lumpH+3+lumpD+lumpT, group_label = "Gas Transmission")%>%
  cat(., file = paste0(exportdir, "table.html"))
```


```{r operators releases}
df.release <- df %>%
  filter(MSYS == "HL") %>%
  mutate( TOTAL_RELEASE = scale(TOTAL_RELEASE) ) %>%
  rbind(  
    filter(df, MSYS == "Gas") %>% mutate(TOTAL_RELEASE = scale(TOTAL_RELEASE))
    ) %>%
  mutate(SIGSER = if_else(SIGNIFICANT == "YES", 
                          if_else(SERIOUS == "YES", 2,1),
                          0),
         pri.name = if_else(grepl("CENTERPOINT", pri.name), 
                            paste0("Centerpoint (", SYS, ")"),
                            pri.name))

df.release %>%
  ggplot(aes(x = MDY, y = TOTAL_RELEASE, size = TOTAL_COST_CURRENT/1000000, color = as.factor(SIGSER)))+
  geom_point(alpha = .9)+
  #adjust labeling based on whatever 
  geom_label_repel(data = filter(df.release, TOTAL_RELEASE > 3 | SIGSER > 1 & FATAL >0),
                  aes(x = MDY, y = TOTAL_RELEASE, label = pri.name),
                  size = 4, color = "gray 20",  parse = F,
                  box.padding = .6, min.segment.length = .2, 
                  nudge_y = -0.2, nudge_x = 15, seed = 2020,)+
 # annotate("text", x = ymd("2010-12-30"), y = 8, label = "ok")+
  theme_pst_lite()+
  theme(legend.position = "right",
        legend.direction = "vertical")+
  scale_color_brewer(palette = "Purples", name = "Significance",
                     labels = c("None", "Significant","Serious"))+
  scale_y_continuous(limits = c(-1.1,6))+
  scale_size(breaks = seq(0,100,25), 
             labels =scales::unit_format(prefix = "$",
                                         suffix = "M") ,
             name = "Damage Cost\n(2022 USD)")+
  labs(caption = paste0("Release Z-Scores are calculated to unit of system measurement:\n MSCF for Natural Gas and BBLs for Liquids." ,"\n",capt),
       tag = tag,
       x = "Date",
       y = "Release Z-Score",
       title = paste0(moStateL," Pipeline Incident Detail"),
       subtitle = "Total Release Z-Scores, Damage Cost, and Significance (2010-2021)")

plotName = "OpRelease.png"
plotExport = paste0(exportdir, plotName)
wdt = 2560
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "px")
```


```{r operators by cause tiles}
df.c <- df %>%
  filter(pri.name %in% bigNames | pri.name %in% newNames) %>%
  mutate(pri.name = if_else(grepl(newNames[1], pri.name),
                            newNames[1],
                            if_else(grepl(newNames[2],pri.name),
                                    newNames[2],
                                    pri.name)))%>%
  group_by(pri.name)%>%
  mutate(inc = n())%>%
  ungroup()%>%
  group_by(pri.name, CAUSE) %>%
  summarise(cPer = n()/inc, 
            inc = sum(unique(inc), na.rm=T))%>%
  ungroup()%>%
  complete(pri.name, CAUSE, fill = list(cPer = 0))%>%
  arrange(inc) %>%
  mutate(pri.name = factor(pri.name),
         pri.name = fct_reorder(pri.name, desc(inc)),
         colText = if_else(cPer >.5 , "black", "white"))

df.c%>%
  ggplot(aes(y=str_wrap(pri.name, 15), x = str_wrap(str_to_sentence(CAUSE),15), fill = cPer))+
  geom_tile(width = .8, height = .8)+
  geom_text(aes(label = percent(cPer, accuracy = 1), color = colText, 
                fontface = ifelse(colText == "black", "bold","plain")), 
            size = 6, family = pstFont)+
  scale_fill_viridis(name = "Percent\nof Incidents", labels = percent_format())+
  scale_color_manual(values = c("gray25", "gray75"), guide = NULL)+
  scale_x_discrete(position = "top", name = NULL)+
  labs(y = NULL,
       title = "Incident Cause Proportions by Operator",
       subtitle = paste0("Selected Operators in ", moStateL, " (2010-2021)"),
       tag = tag, 
       caption = capt)+
  theme_pst_lite()+
  theme(axis.text.x = element_text(angle = -36, hjust = 1, size = 6.6),
        axis.text.y = element_text( size = 6.6),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6.6))

plotName = "OpCause.png"
plotExport = paste0(exportdir, plotName)
wdt = 8
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "in")
```


## Spatial Analysis  
```{r load state shapefile}
library(sf)
projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

stateShape <- read_sf("gis/us_states.shp")

stateShape <- stateShape[stateShape$NAME == moStateL,]

#mb_access_token("accesstokenhere", install = T)
```

```{r map, include = TRUE, fig.fullsize = TRUE}
df.sp <- st_as_sf(df, 
                  coords = c("LOCATION_LONGITUDE","LOCATION_LATITUDE"),
                  crs = projcrs) 

df.state <- st_join(df.sp, st_transform(stateShape, crs = projcrs), join = st_within) %>%
  filter(NAME.y == moStateL) %>%
  st_transform(crs = 3857)

bbox <- st_bbox(df.state, crs = 3857)

base_tiles <- get_static_tiles(location = df.state,
                               zoom = 5,
                               style_id = "cl1pbimv7000114ouzmsb6obk",
                               username = "jimeagah",
                               buffer_dist = 1500000,
                               scaling_factor = "1x",
                               crop = T)

place_tiles <- get_static_tiles(location = stateShape,
                                zoom = 5,
                                style_id = "cl6zaoczy004n14jqcivrkvb5",
                                username = "jimeagah",
                                buffer_dist = 100,
                                scaling_factor = "2x",
                                crop = F)

```

```{r}
bbox <- st_bbox(df.state, crs = 3857)
xdist <- (bbox[[3]]-bbox[[1]])
bxn <- bbox[[1]] - (.2*xdist)
bxx <- bbox[[3]] + (.1*xdist)
byn <-  mean(c(bbox[[4]],bbox[[2]])) - ((xdist / 1.62)/2)
byx <- mean(c(bbox[[4]],bbox[[2]])) + ((xdist / 1.62)/2)



##ggspatial
df.state%>%
  ggplot()+
  layer_spatial(base_tiles)+
  geom_sf(data = stateShape, fill = NA, color = "gray50")+
  layer_spatial(place_tiles)+
  geom_sf(aes(color = SYS, size = TOTAL_COST_CURRENT), alpha = .6)+
  scale_size(name = "Cost of Damage",
             breaks = seq(0,100000000, 25000000),
             labels = scales::label_dollar(scale = .000001, suffix = "M", accuracy = 1))+
  scale_color_brewer(palette = "Set2", name = "System",
                     labels = c("Gas Distribution","Gas Transmission", "Hazardous Liquids"))+
  coord_sf(crs = 3857,
           xlim = c(bxn, bxx ),
           ylim = c(byn , byx),
           expand = TRUE)+
  labs(title = paste0("Incident Locations and Cost in ", moStateL),
       subtitle = "All Systems (2010 - 2022)",
       x = NULL,
       y = NULL,
       caption = capt,
       tag = tag)+
  theme_void()+
  theme(legend.position = c(0.2,0.2),
        legend.background = element_rect(fill = "#E8EBEA", colour = NA),
        legend.box = "horizontal",
        text = element_text(pstFont),
        plot.margin = unit(c(0,0,0,0), "cm"),
        plot.tag.position = c(.96,0.95),
        plot.tag = element_text(color = "gray 30", size = 8),
        plot.title = element_text(face = "bold", colour = "black", hjust = .015),
        plot.subtitle = element_text(colour = "black", hjust = .015),
        plot.caption = element_text(colour = "gray 30"),
        legend.title = element_text(colour = "gray 10", face = "bold"),
        legend.text = element_text(colour = "gray 10"),
        plot.background = element_rect(fill = "#E8EBEA", colour = NA)) #fill is to match bg

plotName = "Map.png"
plotExport = paste0(exportdir, plotName)
wdt = 8
hgt = wdt/1.62
ggsave(plotExport, width = wdt, height = hgt, units = "in")
``` 


