---
title: "Untitled Analysis"
subtitle: "Specifics (Timeline)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout: default
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE, echo = FALSE}
rm(list = ls())
```


```{r setup, include=FALSE}
library(tufte)
library(incays)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(sp)
library(sf)
library(rgeos)
library(tufte)
library(rgdal)
library(ggspatial)
library(mapboxapi)

knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE)
options(knitr.table.format = "html")
capt = "Source: PHMSA Incident and Mileage Data (2010-2021)"
tag = "PST 2022"
##
moState = "OH"
moStateL = state.name[match(moState,state.abb)]
```

```{r theme_pst}
library(showtext)
#> Loading required package: sysfonts
#> Loading required package: showtextdb
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()
theme_pst <- function(baseSize=8) {
  (theme(
                       axis.line.x = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.line.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.ticks.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       text = element_text(family = pstFont),
                       axis.text = element_text(colour = "#8C9394",
                                                size = baseSize * .8),
                       axis.title = element_text(colour ="black"),
                       panel.grid.minor.y = element_line(linetype = "dotted", 
                                                         colour = "#C4C8C6"),
                       panel.grid.major.y = element_line(colour = "#394C56", 
                                                         linetype = "dotted"),
                       panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(),
                       panel.background = element_blank(),
                       panel.border = element_blank(),
                       strip.background = element_rect(linetype = 0),
                       strip.text = element_blank(),
                       strip.text.x = element_blank(),
                       strip.text.y = element_blank(),
                       plot.tag.position = "topright",
                       plot.tag = element_text(size = baseSize *.66,
                                               colour = "#8C9394"),
                       legend.text = element_text(colour = "black",
                                                  size = baseSize * .8),
                       legend.background = element_rect(fill = NULL, 
                                                        color = "#182125"),
                       legend.title = element_text( face = "bold", 
                                                    colour = "black",
                                                    size = baseSize),
                       legend.position = "right",
                       legend.key = element_blank(),
                       #legend.background = element_blank(),
                       plot.background = element_rect(fill = "#ffffff"),
                       plot.title = element_text(face = "bold", 
                                                 colour = "black",
                                                 size = baseSize*1.2),
                       plot.subtitle = element_text(colour = "#8C9394",
                                                    size = baseSize)
                     )
  )
}
```

```{r nice colors}
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]
```

```{r loading incident data}
#all incident data + op mileage
df <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv") %>%
  filter(STATE == moState)%>%
  mutate(MSYS = str_sub(SYSTEM_TYPE, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

adj.states <- c("PA","WV","KY","MI","IN")
df.adj <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv")%>%
  filter(STATE %in% adj.states)%>%
  mutate(MSYS = str_sub(SYSTEM_TYPE, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

df.all <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv")%>%
  filter(STATE %in% state.abb & STATE != moState)%>%
  mutate(MSYS = str_sub(SYSTEM_TYPE, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

OHmiles <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv") %>%
  filter(STATE == moState,
         IYEAR >= 2010)%>%
  mutate(SYSTEM_TYPE = str_sub(SYSTEM_TYPE, 1,2))%>%
  group_by(IYEAR, SYSTEM_TYPE)%>%
  summarise(mileage= sum(mileage, na.rm=T))

miles <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv") %>%
  filter(IYEAR == 2021)%>%
  group_by(STATE)%>%
  summarise(mileage= sum(mileage, na.rm=T))


newMiles <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv")%>%
  filter(IYEAR == 2021, STATE == moState)

newSM <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv") %>% 
  filter(IYEAR >=2010,
         STATE == moState | STATE %in% adj.states)%>%
  mutate(STATE = if_else(STATE != moState, "Adjacent (avg.)", STATE))%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  mutate(mileage = if_else(STATE != moState, mileage/length(adj.states), mileage))

allSM <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv") %>% 
  filter(IYEAR >=2010, STATE != moState)%>%
  mutate(STATE = "Rest of U.S. (avg.)")%>%
  group_by(STATE, IYEAR)%>%
  summarise(mileage = sum(mileage)) %>%
  mutate(mileage = if_else(STATE != moState, mileage/length(unique(df.all$STATE)), mileage)) %>%
  rbind(newSM)
```

# Pipeline Incident Analysis  

\newthought{This makes it caps} or something to start new sections. Links might not show up properly so use this formatting: \href{https://www.phmsa.dot.gov/data-and-statistics/pipeline/pipeline-incident-flagged-files}{\textcolor{Cerulean}{\underline{PHMSA's website}}}. 

## Overview  
```{r quick facts}
sig <- df %>%
  count(SIGNIFICANT)%>%
  filter(SIGNIFICANT == "YES")

bad <- df %>%
  count(INJURY_IND, FATALITY_IND) %>%
  filter(INJURY_IND != "NO" | FATALITY_IND != "NO") %>%
  summarise(n = sum(n))

fire <- df %>%
  count(EXPLODE_IND, IGNITE_IND) %>%
  filter(EXPLODE_IND != "NO" | IGNITE_IND != "NO") %>%
  summarise(n = sum(n))

death <- df %>%
  mutate(FATAL = replace_na(FATAL,0)) %>%
  summarise(FATAL = sum(FATAL))

injure <- df %>%
  mutate(INJURE = replace_na(INJURE,0)) %>%
  summarise(INJURE = sum(INJURE))

evac <- df %>%
  mutate(NUM_PUB_EVACUATED = replace_na(NUM_PUB_EVACUATED,0)) %>%
  summarise(NUM_PUB_EVACUATED = sum(NUM_PUB_EVACUATED))

rel <- df %>%
  mutate(UNINTENTIONAL_RELEASE = replace_na(UNINTENTIONAL_RELEASE,0),
         INTENTIONAL_RELEASE = replace_na(INTENTIONAL_RELEASE,0),
         TOTAL_RELEASE = UNINTENTIONAL_RELEASE + INTENTIONAL_RELEASE) %>%
  group_by(SYSTEM_TYPE)%>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

gasRel <- filter(rel, !grepl("HL", SYSTEM_TYPE)) %>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

hazRel <- filter(rel, grepl("HL", SYSTEM_TYPE))

modeFind <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

freqOp <- modeFind(df$NAME)
```
**Quick Facts**  
  
From 2010 to today, `r moStateL` has had:  
  
*  `r nrow(df)` total incidents including:  

  +  `r nrow(filter(df,grepl("HL",SYSTEM_TYPE)))` Hazardous Liquid incidents  
  +  `r nrow(filter(df,grepl("GT",SYSTEM_TYPE)))` Gas Transmission incidents    
  +  `r nrow(filter(df,grepl("GD",SYSTEM_TYPE)))` Gas Distribution incidents 
  
*  `r sig[[2]]` significant incidents
*  `r bad[[1]]` incidents ending in injury and/or fatality
*  `r fire[[1]]` incidents resulting in fire and/or explosion
*  `r comma(evac[[1]])` total members of the public evacuated 
*  Total releases of:  

  +  `r comma(gasRel[[1]])` Thousand Cubic Feet (mscf) of gas  
  +  `r comma(hazRel[[2]]*42)` gallons of hazardous liquids

```{marginfigure include = T, echo = T}
Place margin text like this 
```

## Incident Breakdown

```{r}
df %>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(df.adj, IYEAR), n = n/6, STATE = "Adjacent (avg.)"),
    mutate(count(df.all, IYEAR), n = n/length(unique(df.all$STATE)), STATE = "Rest of U.S. (avg.)")
    ) %>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/mileage)%>%
  ggplot(aes(x = IYEAR, y = n, group = STATE, color = STATE))+
  geom_point(alpha = .8)+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State(s)",
                     breaks = c(moState, "Adjacent (avg.)", "Rest of U.S. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "Rest of U.S. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(5,23),
                    breaks = seq(5,25,5))+ 
  theme(legend.position = c(.15, .825),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Pipeline Incident Rate"),
       subtitle = "All Systems, All Incidents (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Incidents", x = "Year")

wdt = 2560
hgt = wdt/1.62
ggsave("./OH/AllInc.png", width = wdt, height = hgt, units = "px")
```

```{r}
df %>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(df.adj, IYEAR), n = n/6, STATE = "Adjacent (avg.)"),
    mutate(count(df.all, IYEAR), n = n/length(unique(df.all$STATE)), STATE = "Rest of U.S. (avg.)")
    ) %>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/mileage)%>%
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  geom_point(aes(size = n), alpha = .8)+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "Rest of U.S. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "Rest of U.S. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(0.00005,.00048),
                     breaks = seq(.0001,.0004, .0001), labels = c(1,2,3,4))+ 
  theme(legend.position = c(.8, .88),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Pipeline Incident Rate"),
       subtitle = "All Systems, All Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Incident Count")

wdt = 2560
hgt = wdt/1.62
ggsave("./OH/AllIncMile.png", width = wdt, height = hgt, units = "px")
```

```{r}
df %>%
  filter(SIGNIFICANT == "YES")%>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(filter(df.adj, SIGNIFICANT == "YES"), IYEAR), n = n/6, STATE = "Adjacent (avg.)"),
    mutate(count(filter(df.all, SIGNIFICANT == "YES"), IYEAR), n = n/length(unique(df.all$STATE)), STATE = "Rest of U.S. (avg.)")
    )%>%
  filter(IYEAR < 2022)%>%
  left_join(allSM) %>% 
  mutate(ir = n/mileage) %>% 
  ggplot(aes(x = IYEAR, y = ir, group = STATE, color = STATE))+
  geom_point(aes(size = n), alpha = .8)+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State",
                     breaks = c(moState, "Adjacent (avg.)", "Rest of U.S. (avg.)"),
                     labels = c(moStateL, "Adjacent (avg.)", "Rest of U.S. (avg.)"))+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(0.000025,.00028),breaks = seq(.00005,.00025, .00005), labels = c(0.5,1,1.5,2, 2.5))+ 
  theme(legend.position = c(.15, .825),
        legend.box = "horizontal")+
  labs(title = paste(moStateL, "Significant Incident Rate"),
       subtitle = "All Systems, Significant Incidents per 1,000 Miles (2010-2021)",
       caption = capt,
       tag = tag,
       y = "Incident Rate (per 1,000 mi)", 
       x = "Year", 
       size = "Incident Count")

wdt = 2560
hgt = wdt/1.62
ggsave("./OH/SigIncMile.png", width = wdt, height = hgt, units = "px")
```


```{r}


  
```


## Operator Analysis 

```{r}
op <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/operator_id_match.csv") %>%
  select(!`...1`)


df.ops <- df %>%
  mutate(sys = gsub(".*\\((.*)\\).*", "\\1", SYSTEM_TYPE)) %>%
  left_join(op, by = c("OPERATOR_ID" = "sub.id"))%>% 
  distinct(NARRATIVE, .keep_all = T)
  # mutate(pri.name = str_to_upper(pri.name),
  #        pri.name = str_replace(pri.name, 
  #                               "ONEOK FIELD SERVICES COMPANY, L.L.C.", 
  #                               "ONEOK"),
  #        pri.name = str_replace(pri.name, 
  #                               "PLAINS MARKETING, L.P.", 
  #                               "PLAINS"),
  #        pri.name = str_replace(pri.name, 
  #                               "BKEP CRUDE", 
  #                               "BKEP PIPELINE"))%>%
  # count(pri.name, sub.name)%>%
  # group_by(pri.name)%>%
  # summarise(n = n())
```


```{r}
df.t <- df.ops %>%
  mutate(SIGNIFICANT = if_else(SIGNIFICANT == "YES", 1,0),
         EXPLODE_IND = if_else(EXPLODE_IND == "YES", 1,0),
         IGNITE_IND = if_else(IGNITE_IND == "YES", 1,0),
         SYSTEM_TYPE = str_sub(SYSTEM_TYPE, 1,2)) %>%
  group_by(pri.name, SYSTEM_TYPE)%>%
  summarise(miles = mean(mileage),
            n = n(),
            sig = sum(SIGNIFICANT)/n,
            ipm = n/(miles/1000),
            trend = "",
            fatal = sum(FATAL),
            injure = sum(INJURE),
            evac = sum(NUM_PUB_EVACUATED, na.rm = T),
            cost = sum(TOTAL_COST_CURRENT)/1000,
            cpi = cost/ n ,
            rel = sum(TOTAL_RELEASE),
            rpi = rel/n ) %>%
  ungroup()

t.gd <- df.t %>%
  filter(SYSTEM_TYPE == "GD") %>%
  mutate(pri.name = if_else(pri.name == "Questar", "Questar (GD)", pri.name),
         pri.name = fct_lump(f=pri.name, 3, w=n, other_level = "Others (GD, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

t.gt <- df.t %>%
  filter(SYSTEM_TYPE == "GT") %>%
  mutate(pri.name = if_else(pri.name == "Questar", "Questar (GT)", pri.name),
         pri.name = fct_lump(f=pri.name, 3, w=n, other_level = "Others (GT, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T) %>%
  arrange(desc(n))

t.hl <- df.t %>%
  filter(SYSTEM_TYPE == "HL") %>%
  mutate(pri.name = fct_lump(f=pri.name, 3, w=n, other_level = "Others (HL, avg.)"))%>%
  group_by(pri.name)%>%
  summarise_at(vars(n:rpi), mean, na.rm = T)%>%
  arrange(desc(n))

bigNames <- as.character(c(t.gd$pri.name, t.gt$pri.name,t.hl$pri.name))
bigNames <- bigNames[!grepl("Others",bigNames)]

namesMatch <- tibble(pri.name = bigNames, MSYS = c(rep("GD",3),rep("GT",3),rep("HL",3)))

df.trend <- df.ops %>%
  filter(IYEAR <2022,
         MSYS != "UNGS",
         MSYS != "GG")%>%
  left_join(namesMatch, by = c("pri.name", "MSYS"), keep = T) %>%
  mutate(pri.name.y = if_else(pri.name.x == "Questar",
                              if_else(MSYS.x == "GT", "Questar (GT)", "Questar (GD)"),
                              pri.name.y),
         pri.name.y = if_else(is.na(pri.name.y),
                              if_else(MSYS.x == "GT", "Others (GT Avg.)",
                                      if_else(MSYS.x == "GD", "Others (GD Avg.)", 
                                              if_else(MSYS.x == "HL", "Others (HL Avg.)",
                                                      "N"))),
                              pri.name.y),
         MSYS.y = coalesce(MSYS.y, MSYS.x))%>%
  rename(pri.name = pri.name.y,
         MSYS = MSYS.y)%>%
  count(pri.name, MSYS, IYEAR) %>%
  complete(pri.name, MSYS, IYEAR, fill = list(n = 0)) %>%
  left_join(namesMatch, by = c("pri.name", "MSYS"), keep = T) %>%
  filter(!is.na(pri.name.y) | grepl("Others",pri.name.x))%>%
  mutate(drop = if_else(!str_detect(pri.name.x,MSYS.x) & 
                          (grepl("Others",pri.name.x)| grepl("Questar", pri.name.x)), T, F))%>%
  filter(!drop)%>%
  rename(pri.name = pri.name.y)%>%
  group_by(pri.name)%>%
  mutate(tot = sum(n),
         n = (n/tot)*100,
         pri.name = if_else(is.na(pri.name), pri.name.x, pri.name))

df.lines <- split(df.trend$n, df.trend$pri.name)


df.k <- rbind(t.gd, t.gt, t.hl)

df.k %>%
  select(!c(fatal, injure,evac)) %>%
  mutate(trend = "")%>%
  mutate(n = round(n,0),
         sig = paste0(round(sig*100, 0), "%"))%>%
  mutate_at(vars(cost:rpi), round, 0)%>%
  kbl(col.names = c("Parent Operator", "Total", "Percent Significant",
                    "per 1K Miles", "Trend","Total","per Incident", "Total","per Incident"),
      format.args = list(big.mark = ","), digits = 2, booktabs = TRUE)%>%
  kable_classic_2(lightable_options = c("striped","hover"))%>%
  add_header_above(c(" "=1, "Incidents" = 4, "Cost\n(K, USD)" =2, "Releases" = 2)) %>%
  column_spec(5, image = spec_plot(df.lines, same_lim = T, 
                                   min = list(col = orange, pch = 20, cex = .2),
                                   max = list(col = green, pch = 20, cex = .2)))%>%
  column_spec(4, color=spec_color(df.k$ipm, option = "B", end = .7), bold = T)%>%
  column_spec(7, color=spec_color(df.k$cpi, option = "B", end = .7), bold = T)%>%
  column_spec(9, color=spec_color(df.k$rpi, option = "B", end = .7), bold = T) %>%
  group_rows(start_row = 1, end_row = 4, group_label = "Gas Distribution")%>%
  group_rows(start_row = 5, end_row = 8, group_label = "Gas Transmission")%>%
  group_rows(start_row = 9, end_row = 12, group_label = "Hazardous Liquids")%>%
  cat(., file = "./OH/table.html")

  




```


```{r}

df.ops %>%
  left_join(namesMatch, keep = T) %>% 
  mutate(pri.name.y = if_else(pri.name.x == "Questar", "Questar",
                              if_else(is.na(pri.name.y), "Others (Avg.)",
                                      pri.name.y)),
         ghl = if_else(MSYS.x == "HL","HL","Gas"),
         pri.name.f = factor(pri.name.y,
                             levels = c(namesMatch$pri.name, "Questar","Others (avg.)")),
         pri.name.f = replace_na(pri.name.f, "Others (avg.)")
         ) %>% 
  ggplot(aes(x = TOTAL_RELEASE, y = TOTAL_COST_CURRENT, color = pri.name.f))+
  geom_point(alpha = .75)+
  scale_x_continuous(trans = "pseudo_log", 
                     breaks = c(0,50,5000,500000),
                     labels = c("0","50","5K","500K"))+
  scale_y_continuous(trans = "pseudo_log", 
                     breaks = c(0,100,10000,1000000,100000000),
                     labels = c("0","100","10K","1M","100M"))+
  facet_wrap(~ghl, scales = "free")+
  scale_color_brewer(palette = "Set1", name = "Operator")+
  labs(title = paste0(moStateL, " Pipeline Incidents"),
       subtitle = "All Systems (2010 - 2021)",
       y = "Cost of Damage (2022 USD)",
       x = "Release Size (Gas: mscf | Haz. Liq.: BBL)",
       tag = tag,
       caption = capt)+
  theme_pst()+
  theme(strip.text.x = element_text(size = 8),
        strip.background = element_blank(),
        legend.position = c(.86,.305),
        legend.background = element_rect(fill = rgb(1,1,1,.6)))
  
wdt = 2560
hgt = wdt/1.62
ggsave("./OH/RelOp.png", width = wdt, height = hgt, units = "px")
```


## Spatial Analysis  
```{r load state shapefile}
projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

stateShape <- read_sf("/Users/jameseager/Documents/projects/PHMSA_clean/data/gis/us_states.shp")

stateShape <- stateShape[stateShape$NAME == moStateL,]

#mb_access_token("pk.eyJ1IjoiamltZWFnYWgiLCJhIjoiY2l2aWw3YmJnMDIxdzJ5bXFuZDZ4bnU1YiJ9.lMwViZI1fDl75hf6XGm78g", install = T)
```

```{r map, include = TRUE, fig.fullsize = TRUE}
df.sp <- st_as_sf(df, 
                  coords = c("LOCATION_LONGITUDE","LOCATION_LATITUDE"),
                  crs = projcrs) 

df.state <- st_join(df.sp, st_transform(stateShape, crs = projcrs), join = st_within) %>%
  filter(NAME.y == moStateL)

pst_tiles <- get_static_tiles(location = stateShape,
                              zoom = 5,
                              style_id = "cl1pbimv7000114ouzmsb6obk",
                              username = "jimeagah",
                              buffer_dist = 1500000,
                              crop = T)

## tag 
place_label <- function(label, size = 6, ...) {
  annotate("text", label = label, x = Inf, y = Inf, 
           hjust = 0, vjust = 1, size = size, ...)
}
##ggspatial
df.state%>%
  ggplot()+
  layer_spatial(pst_tiles)+
  geom_sf(data = stateShape, fill = NA, color = "black")+
  geom_sf(aes(color = MSYS, size = TOTAL_COST_CURRENT), alpha = .6)+
  scale_color_brewer(palette = "Set2", name = "System",
                     labels = c("Gas Distribution","Gas Gathering","Gas Transmission",
                                "Hazardous Liquids","Underground Natural Gas Storage"))+
  scale_size(name = "Cost of Damage\n(Mil. USD)",
             breaks = seq(0,20000000, 5000000),
             labels = seq(0,20,5))+
  coord_sf(crs = 3857,
           xlim = c(-9945541.519513 , -8495684.65),
           ylim = c(4553839.430206, 5321011.87))+
  labs(title = paste0("Incident Locations in ", moStateL),
       subtitle = "All Systems (2010 - Present)",
       x = NULL,
       y = NULL,
       tag = tag,
       caption = capt)+
  theme_void()+
  theme(legend.position = c(0.15,0.42),
        text = element_text(pstFont),
        plot.tag.position = c(.96,0.05),
        plot.tag = element_text(color = "gray 90", size = 8),
        plot.title = element_text(face = "bold", colour = "#ffffff", hjust = .015),
        plot.subtitle = element_text(colour = "#ffffff", hjust = .015),
        plot.caption = element_text(colour = "gray 90"),
        legend.title = element_text(colour = "gray 90", face = "bold"),
        legend.text = element_text(colour = "gray90"),
        plot.background = element_rect(fill = "#457969", colour = NA))

wdt = 2560
hgt = wdt/1.62
ggsave("./OH/MapInc.png", width = wdt, height = hgt, units = "px")
```


## Sources  
  
**Data:**  


**Tech:**  
