---
title: "State of Safety"
subtitle: "Oklahoma (2010-2021)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout:
    latex_engine: xelatex
    keep_tex: true
  tufte::tufte_html: default
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE, echo = FALSE}
rm(list = ls())
```


```{r setup, include=FALSE}
library(tufte)
library(incays)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

library(tidyverse)
library(ggrepel)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(sp)
library(sf)
library(rgeos)
library(tufte)
library(rgdal)
library(ggspatial)
library(mapboxapi)


knitr::opts_chunk$set(echo = FALSE, 
                      include = FALSE, 
                      message = FALSE)
options(knitr.table.format = "latex")
capt = "Source: PHMSA Incident and Mileage Data (2010-2021)"

##
moState = "OK"
moStateL = state.name[match(moState,state.abb)]
```

```{r theme_pst}
library(showtext)
#> Loading required package: sysfonts
#> Loading required package: showtextdb
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()
theme_pst <- function(baseSize=10) {
  (theme(
                       axis.line.x = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.line.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.ticks.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       text = element_text(family = pstFont),
                       axis.text = element_text(colour = "#8C9394",
                                                size = baseSize * .8),
                       axis.title = element_text(colour ="black"),
                       panel.grid.minor.y = element_line(linetype = "dotted", 
                                                         colour = "#C4C8C6"),
                       panel.grid.major.y = element_line(colour = "#394C56", 
                                                         linetype = "dotted"),
                       panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(),
                       panel.background = element_rect(fill = "#ffffff"),
                       panel.border = element_blank(),
                       strip.background = element_blank(),
                       strip.text = element_blank(),
                       strip.text.x = element_text(size = baseSize*.8, 
                                                   face = "bold"),
                       strip.text.y = element_blank(),
                       legend.text = element_text(colour = "black",
                                                  size = baseSize * .8),
                       legend.background = element_rect(fill = NULL, 
                                                        color = "#182125"),
                       legend.title = element_text( face = "bold", 
                                                    colour = "black",
                                                    size = baseSize),
                       legend.position = "right",
                       legend.key = element_blank(),
                       #legend.background = element_blank(),
                       plot.background = element_rect(fill = "#ffffff"),
                       plot.title = element_text(face = "bold", 
                                                 colour = "black",
                                                 size = baseSize*1.2),
                       plot.subtitle = element_text(colour = "#8C9394",
                                                    size = baseSize)
                     )
  )
}
```

```{r nice colors}
col <- brewer.pal(6,"Set2")
green <- col[1]
orange <- col[2]
purple <- col[3]
pink <- col[4]
lime <- col[5]
yellow <- col[6]
```

```{r loading incident data}
#all incident data + op mileage
df <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv") %>%
  filter(STATE == moState)%>%
  mutate(MSYS = str_sub(SYSTEM_TYPE, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))

adj.states <- c("NM","TX","CO","KS","MO","AR")
df.adj <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv")%>%
  filter(STATE %in% adj.states)%>%
  mutate(MSYS = str_sub(SYSTEM_TYPE, 1,2),
         MSYS = if_else(MSYS == "UN", "UNGS",MSYS))
```

```{r}
miles <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/all_inc.csv") %>%
  distinct(OPERATOR_ID, STATE, .keep_all = T)%>%
  group_by(STATE) %>%
  summarise(miles = sum(mileage))

stateMiles <- filter(miles, STATE == "OK") %>% select(miles)


newMiles <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/sys_miles.csv")%>%
  filter(IYEAR == 2021) %>%
  group_by(STATE) %>%
  summarise(mileage = sum(mileage))
  
```


\newthought{At `r stateMiles[[1]]` miles}, Oklahoma has the third most mileage of operational pipelines in the United States. Let’s take a deeper look at the state of Oklahoma and how their operators perform when it comes to pipeline safety. 

```{r quick facts}
sig <- df %>%
  count(SIGNIFICANT)%>%
  filter(SIGNIFICANT == "YES")

bad <- df %>%
  count(INJURY_IND, FATALITY_IND) %>%
  filter(INJURY_IND != "NO" | FATALITY_IND != "NO") %>%
  summarise(n = sum(n))

fire <- df %>%
  count(EXPLODE_IND, IGNITE_IND) %>%
  filter(EXPLODE_IND != "NO" | IGNITE_IND != "NO") %>%
  summarise(n = sum(n))

death <- df %>%
  mutate(FATAL = replace_na(FATAL,0)) %>%
  summarise(FATAL = sum(FATAL))

injure <- df %>%
  mutate(INJURE = replace_na(INJURE,0)) %>%
  summarise(INJURE = sum(INJURE))

evac <- df %>%
  mutate(NUM_PUB_EVACUATED = replace_na(NUM_PUB_EVACUATED,0)) %>%
  summarise(NUM_PUB_EVACUATED = sum(NUM_PUB_EVACUATED))

rel <- df %>%
  mutate(UNINTENTIONAL_RELEASE = replace_na(UNINTENTIONAL_RELEASE,0),
         INTENTIONAL_RELEASE = replace_na(INTENTIONAL_RELEASE,0),
         TOTAL_RELEASE = UNINTENTIONAL_RELEASE + INTENTIONAL_RELEASE) %>%
  group_by(SYSTEM_TYPE)%>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

gasRel <- filter(rel, !grepl("HL", SYSTEM_TYPE)) %>%
  summarise(TOTAL_RELEASE = sum(TOTAL_RELEASE))

hazRel <- filter(rel, grepl("HL", SYSTEM_TYPE))

modeFind <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

freqOp <- modeFind(df$NAME)
```


**Quick Facts**  
  
From 2010 to today, `r moStateL` has had:  
  
*  `r nrow(df)` total incidents including:  

  +  `r nrow(filter(df,grepl("HL",SYSTEM_TYPE)))` Hazardous Liquid incidents  
  +  `r nrow(filter(df,grepl("GT",SYSTEM_TYPE)))` Gas Transmission incidents    
  +  `r nrow(filter(df,grepl("GD",SYSTEM_TYPE)))` Gas Distribution incidents 
  
*  `r sig[[2]]` significant incidents
*  `r bad[[1]]` incidents ending in injury and/or fatality
*  `r fire[[1]]` incidents resulting in fire and/or explosion
*  `r comma(evac[[1]])` total members of the public evacuated 
*  Total releases of:  

  +  `r comma(gasRel[[1]])` Thousand Cubic Feet (mscf) of gas  
  +  `r comma(hazRel[[2]]*42)` gallons of hazardous liquids

```{marginfigure include = F, echo = F}
Place margin text like this 
```

## Incident Breakdown

```{r}
df %>%
  count(IYEAR, SIGNIFICANT) %>%
  filter(SIGNIFICANT == "YES") %>%
  bind_rows( df %>% count(IYEAR) %>% mutate(SIGNIFICANT = "ALL")) %>%
  ggplot(aes(x = IYEAR, y = n, group = SIGNIFICANT, color = SIGNIFICANT))+
  geom_point()+
  geom_line()+
  theme_clean()+
  scale_color_brewer(palette = "Set2", labels = c("All","Significant"),
                     name = "Type")+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  theme(legend.position = c(.11, .8))+
  labs(title = paste(moStateL, " Pipeline Incidents"),
       caption = capt,
       y = "Incidents", x = "Year")
```

```{r include = TRUE, fig.fullsize = TRUE}
df %>%
  count( STATE, IYEAR) %>%
  rbind(
    mutate(count(df.adj, IYEAR), n = n/6, STATE = "Adjacent (avg.)")
    )%>%
  ggplot(aes(x = IYEAR, y = n, group = STATE, color = STATE))+
  geom_point()+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State")+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(20,60))+
  theme(legend.position = c(.125, .85))+
  labs(title = paste(moStateL, "Pipeline Incidents"),
       subtitle = "All Systems, All Incidents (2010-2021)",
       caption = capt,
       y = "Incidents", x = "Year")

wdt = 2560
hgt = wdt/1.62
ggsave("./OK/AllInc.png", width = wdt, height = hgt, units = "px")
```

```{r}
df %>%
  count( STATE, IYEAR, MSYS) %>%
  rbind(
    mutate(count(df.adj, IYEAR, MSYS), n = n/6, STATE = "Adjacent (avg.)")
    )%>%
  filter(MSYS %in% c("HL","GD","GT"))%>%
  complete(IYEAR, MSYS, STATE, fill = list(n = 0))%>%
  ggplot(aes(x = IYEAR, y = n, group = STATE, color = STATE))+
  geom_point()+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State")+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  #scale_y_continuous(limits = c(20,60))+
  theme(legend.position = c(.125, .85))+
  facet_wrap(~MSYS)+
  labs(title = paste(moStateL, "Pipeline Incidents"),
       subtitle = "All Systems, All Incidents (2010-2021)",
       caption = capt,
       y = "Incidents", x = "Year")

wdt = 2560
hgt = wdt/1.62
ggsave("./OK/SysInc.png", width = wdt, height = hgt, units = "px")
```

```{r include = TRUE, fig.fullsize = TRUE}
df %>%
  count( STATE, IYEAR, SIGNIFICANT) %>%
  rbind(
    mutate(count(df.adj, IYEAR, SIGNIFICANT), n = n/6, STATE = "Adjacent (avg.)")
    )%>%
  filter(SIGNIFICANT == "YES")%>%
  ggplot(aes(x = IYEAR, y = n, group = STATE, color = STATE))+
  geom_point()+
  geom_line()+
  theme_pst()+
  scale_color_brewer(palette = "Set2",
                     name = "State")+
  scale_x_continuous(labels = seq(2010,2021,2), breaks = seq(2010,2021,2))+
  scale_y_continuous(limits = c(5,25))+
  theme(legend.position = c(.125, .15))+
  labs(title = paste(moStateL, "Pipeline Incidents"),
       subtitle = "All Systems, Significant Incidents (2010-2021)",
       caption = capt,
       y = "Incidents", x = "Year")
wdt = 2560
hgt = wdt/1.62
ggsave("./OK/SigInc.png", width = wdt, height = hgt, units = "px")
```



## Operator Analysis 


```{r}
op <- read_csv("/Users/jameseager/Documents/projects/PHMSA_clean/data/operator_id_match.csv") %>%
  select(!`...1`)


state.ops <- df %>%
  mutate(sys = gsub(".*\\((.*)\\).*", "\\1", SYSTEM_TYPE)) %>%
  left_join(op, by = c("OPERATOR_ID" = "sub.id"))%>% 
  distinct(NARRATIVE, .keep_all = T)%>%
  mutate(pri.name = str_to_upper(pri.name),
         pri.name = str_replace(pri.name, 
                                "ONEOK FIELD SERVICES COMPANY, L.L.C.", 
                                "ONEOK"),
         pri.name = str_replace(pri.name, 
                                "PLAINS MARKETING, L.P.", 
                                "PLAINS"),
         pri.name = str_replace(pri.name, 
                                "BKEP CRUDE", 
                                "BKEP PIPELINE"))%>%
  count(pri.name, sub.name)%>%
  group_by(pri.name)%>%
  summarise(sub.name = n())%>%
  filter(sub.name > 1)
  
```
There are 71 unique operators working in `r moStateL`, per PHMSA's incident reports. There are 45 parent companies involved in running these 71 operators, and 10 of these parent companies are responsible for more than 1 operator: `r unique(state.ops$pri.name)`. These numbers come from PHMSA's safety program data, so may not account for partial ownership on pipelines, but allows us to get a sense for which major operators are active in certain states.  
  
  
Of the `r nrow(df)` incidents from 2010 until 2021,  

```{r include = TRUE}
df.t <- df %>%
  mutate(sys = gsub(".*\\((.*)\\).*", "\\1", SYSTEM_TYPE)) %>%
  left_join(op, by = c("OPERATOR_ID" = "sub.id"))%>% 
  distinct(NARRATIVE, .keep_all = T)%>%
  mutate(pri.name = str_to_upper(pri.name),
         pri.name = str_replace(pri.name, 
                                "ONEOK FIELD SERVICES COMPANY, L.L.C.", 
                                "ONEOK"),
         pri.name = str_replace(pri.name, 
                                "PLAINS MARKETING, L.P.", 
                                "PLAINS"),
         pri.name = str_replace(pri.name, 
                                "BKEP CRUDE", 
                                "BKEP PIPELINE"))%>%
  group_by(pri.name)%>%
  mutate(n = n())%>%
  ungroup()

df.trend  <- df.t %>%
  count(fct_lump(pri.name, 5, w = n), IYEAR)%>%
  rename(NAME = 1)%>%
  group_by(NAME)%>%
  mutate(tot = sum(n),
         n = (n/tot)*100)

df.lines <- split(df.trend$n, df.trend$NAME)

df.release <- df.t %>%
  group_by(MSYS)%>%
  mutate(MRelease = max(TOTAL_RELEASE))%>%
  ungroup()%>%
  group_by(fct_lump(pri.name, 5, w = n), IYEAR)%>%
  summarise(release = TOTAL_RELEASE/MRelease)%>%
  summarise(release = sum(release))%>%
  rename(NAME = 1)

df.box <- split(df.release$release, df.release$NAME)
bigs <- as.character(unique(df.release$NAME)[1:5])

df.other <- df.t%>%
  group_by(fct_lump(pri.name, 5, w = n))%>%
  summarise(Miles = sum(unique(mileage))/40,
            n = n()/40,
            ipm = n/(Miles/1000),
            Sig = sum(grepl("YES", SIGNIFICANT))/40,
            Cost = (sum(TOTAL_COST_CURRENT, na.rm = T)/1000)/40,
            cpi = Cost/n)%>%
  mutate(trend = "",
         rels = "")%>%
  rename(NAME = 1)%>%
  select(NAME, Miles, n, ipm, Sig, trend, Cost, cpi, rels)%>%
  slice(6)%>%
  mutate(NAME = "Other (Avg)")

df.k <- df.t %>%
  group_by(fct_lump(pri.name, 5, w = n))%>%
  summarise(Miles = mean(mileage),
            n = n(),
            ipm = n/(Miles/1000),
            Sig = sum(grepl("YES", SIGNIFICANT)),
            Cost = sum(TOTAL_COST_CURRENT, na.rm = T)/1000,
            cpi = Cost/n)%>%
  mutate(trend = "",
         rels = "")%>%
  rename(NAME = 1)%>%
  mutate(NAME = str_replace(NAME, "Other", "Other (sum)"))%>%
  filter(!grepl("Other",NAME))%>%
  rbind(df.other)%>%  
  select(NAME, Miles, n, ipm, Sig, trend, Cost, cpi)

averager <- function(most, tot, n){
  x = (tot - most) / n
  return(x)
}

kMiles <- df.t %>%
  filter(IYEAR == 2021) %>%
  distinct(OPERATOR_ID, .keep_all = T)%>%
  group_by(fct_lump(pri.name, 5, w = n))%>%
  summarise(miles = sum(mileage))%>%
  rename(pri.name = 1)

kMiles <- kMiles %>%
  mutate(miles = if_else(pri.name == "Other",
                         averager(most = (sum(kMiles$miles)-
                                               kMiles[kMiles$pri.name == "Other",][[2]]),
                                  tot = sum(kMiles$miles),
                                  n = length(unique(df$OPERATOR_ID))-5
                                  ),
                         miles
                         ))




df.k%>%
  kbl(col.names = c("Parent Operator", "OK Pipeline\nMileage", "Total", "per 1K Miles",
                    "Significant", "Trend","Total","per Incident"),
      format.args = list(big.mark = ","), digits = 2, booktabs = TRUE)%>%
  kable_classic_2(lightable_options = c("striped","hover"))%>%
  add_header_above(c(" "=2, "Incidents" = 4, "Cost\n(Thou. USD)" =2))%>%
  column_spec(6, image = spec_plot(df.lines, same_lim = T, 
                                   min = list(col = orange, pch = 20, cex = .2),
                                   max = list(col = green, pch = 20, cex = .2)))%>%
  column_spec(4, color=spec_color(df.k$ipm), bold = T)%>%
  column_spec(8, color=spec_color(df.k$cpi), bold = T)
  cat(., file = "./OK/table.html")
```


```{r}
testing <- df.t %>%
  filter(pri.name == "ENBRIDGE") %>%
  mutates(ipm = weightMiles(pri.name, mileage, IYEAR))


## it gets more complicated: 
#for each operator we have to look up theri mileage in og mileage data 
#then add 0 incident year mileage
#end with weighted averge of 

weightMiles <- function(ops, miles, years){
  x <- tibble(ops, miles) %>%
    group_by(ops, IYEAR)%>%
    summarise(ops = ops
              )
}
```


```{r include = TRUE, fig.fullsize = TRUE}
df.t %>%
  filter(!pri.name%in% bigs)%>%
  mutate(pri.name = "Other")%>%
  rbind(filter(df.t, pri.name%in%bigs))%>%
  mutate(ghl = if_else(MSYS == "HL","HL","Gas"))%>%
  ggplot(aes(x = TOTAL_RELEASE, y = TOTAL_COST_CURRENT, color = pri.name))+
  geom_point(alpha = .6)+
  scale_x_continuous(trans = "pseudo_log", 
                     breaks = c(0,50,5000,500000),
                     labels = c("0","50","5K","500K"))+
  scale_y_continuous(trans = "pseudo_log", 
                     breaks = c(0,100,10000,1000000,100000000),
                     labels = c("0","100","10K","1M","100M"))+
  scale_color_brewer(palette = "Set2",
                     name = "Operator")+
  facet_wrap(~ghl)+
   labs(title = paste0(moStateL, " Pipeline Incidents"),
       subtitle = "All Systems (2010 - Present)",
       y = "Cost of Damage (2022 USD)",
       x = "Release Size (Gas: mscf | Haz. Liq.: BBL)",
       caption = capt)+
  theme_pst()

# df.t %>%
#   filter(TOTAL_COST_CURRENT == max(TOTAL_COST_CURRENT))%>%
#   view()


wdt = 2560
hgt = wdt/1.62
ggsave("./OK/RelInc.png", width = wdt, height = hgt, units = "px")
```

```{r}
df.t %>%
  filter(!pri.name%in% bigs)%>%
  mutate(pri.name = "Other")%>%
  rbind(filter(df.t, pri.name%in%bigs))%>%
  count(IYEAR, pri.name)%>%
  mutate(n = if_else(pri.name == "Other",n/40, as.double(n)))%>%
  ggplot(aes(x = IYEAR, y = n, group = pri.name, color = pri.name))+
  geom_point()+
  geom_line()+
  scale_color_brewer(palette = "Set2")+
  theme_pst()

```


## Spatial Analysis  
```{r load state shapefile}
projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

stateShape <- read_sf("/Users/jameseager/Documents/projects/PHMSA_clean/data/gis/us_states.shp")

stateShape <- stateShape[stateShape$NAME == moStateL,]

#mb_access_token("pk.eyJ1IjoiamltZWFnYWgiLCJhIjoiY2l2aWw3YmJnMDIxdzJ5bXFuZDZ4bnU1YiJ9.lMwViZI1fDl75hf6XGm78g", install = T)
```

```{r map, include = TRUE, fig.fullsize = TRUE}
df.sp <- st_as_sf(df, 
                  coords = c("LOCATION_LONGITUDE","LOCATION_LATITUDE"),
                  crs = projcrs) 

df.state <- st_join(df.sp, st_transform(stateShape, crs = projcrs), join = st_within) %>%
  filter(NAME.y == moStateL)

pst_tiles <- get_static_tiles(location = stateShape,
                              zoom = 5,
                              style_id = "cl1pbimv7000114ouzmsb6obk",
                              username = "jimeagah",
                              buffer_dist = 100000,
                              crop = T)

##ggspatial
df.state%>%
  ggplot()+
  layer_spatial(pst_tiles)+
  geom_sf(data = stateShape, fill = NA, color = "black")+
  geom_sf(aes(color = MSYS), alpha = .6)+
  scale_color_brewer(palette = "Set2", name = "System",
                     labels = c("Gas Distribution","Gas Gathering","Gas Transmission",
                                "Hazardous Liquids","Underground Natural Gas Storage"))+
  coord_sf(crs = 3857,
           xlim = c(-11465907.55, -10519691.88),
           ylim = c(3985303.96, 4439106.79))+
  labs(title = paste0("Incident Locations in ", moStateL),
       subtitle = "All Systems (2010 - Present)",
       x = NULL,
       y = NULL,
       caption = capt)+
  theme_void()+
  theme(legend.position = c(0.15,0.42),
        text = element_text(pstFont),
        plot.title = element_text(face = "bold", colour = "#ffffff"),
        plot.subtitle = element_text(colour = "#ffffff"),
        legend.title = element_text(colour = "gray 90", face = "bold"),
        legend.text = element_text(colour = "gray90"),
        plot.background = element_rect(fill = "#457969", colour = NA))

wdt = 2560
hgt = wdt/1.62
ggsave("./OK/MapInc.png", width = wdt, height = hgt, units = "px")
```




## Sources  
  
**Data:**  


**Tech:**  
