---
title: "NRC Rundown (2 weeks)"
subtitle: "Pipeline Incident Calls (2 weeks)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout: default
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{img/PST_logo.png}\LARGE\\}
  - \posttitle{\end{center}}
editor_options: 
  chunk_output_type: console
---

```{r delete all, include = FALSE, echo = FALSE}
rm(list = ls())
```

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

#relevant libs, some more useful than others
library(tidyverse)
library(ggrepel)
library(lubridate)
library(RColorBrewer)
library(ggthemes)
library(viridis)
library(scales)
library(readxl)
library(kableExtra)
library(ggmap)
library(sp)
library(sf)
library(rgeos)
library(tufte)
library(rgdal)
library(ggspatial)
library(mapboxapi)

#update root directory as needed

#some general knit options
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE, warning = FALSE,
                      kable.force.latex = T)
options(knitr.table.format = "latex")
```

```{r color setup}
colors <- brewer.pal(6,"Set2")
green <- colors[1]
orange <- colors[2]
blue <- colors[3]
pink <- colors[4]
limegreen <- colors[5]
yellow <- colors[6]
```

```{r}
path <- "https://nrc.uscg.mil/FOIAFiles/Current.xlsx"

temp = tempfile(fileext = ".xlsx")
download.file(path, destfile=temp, mode='wb')

all <- read_xlsx(temp, sheet = 2) %>%
  mutate(
    INCIDENT_DATE_TIME = parse_date_time(INCIDENT_DATE_TIME, orders = "mdY HM"),
    INC_DATE = dmy(paste(day(INCIDENT_DATE_TIME), month(INCIDENT_DATE_TIME), year(INCIDENT_DATE_TIME), sep = "-")),
    LOC_FIRST = coalesce(LOCATION_NEAREST_CITY, LOCATION_COUNTY),
    LOC_FULL = paste(LOC_FIRST, LOCATION_STATE, sep = ", ")
    )

commons <- read_xlsx(path, sheet = 2) %>%
  mutate(
    INCIDENT_DATE_TIME = parse_date_time(INCIDENT_DATE_TIME, orders = "mdY HM"),
    INC_DATE = dmy(paste(day(INCIDENT_DATE_TIME), month(INCIDENT_DATE_TIME), year(INCIDENT_DATE_TIME), sep = "-")),
    LOC_FIRST = coalesce(LOCATION_NEAREST_CITY, LOCATION_COUNTY),
    LOC_FULL = paste(LOC_FIRST, LOCATION_STATE, sep = ", ")
    ) %>%
  filter(INC_DATE >= max(INC_DATE) - weeks(2)) %>%
  left_join( read_xlsx(path, sheet = 3), by = "SEQNOS")

#set another text bit for plotting
tr <- stamp("03.22.87")
timerange = paste("From", 
                  tr((max(commons$INC_DATE) - weeks(2))),
                  "to",
                  tr(max(commons$INC_DATE)))
#generic caption text, update to reflect data 
cp <- stamp("8 March, 2022")
capt = paste("Data from NRC FOIA:", timerange)
```


```{r geocoding}
pipes <- commons %>% 
  filter(TYPE_OF_INCIDENT == "PIPELINE") %>%
  mutate(human = if_else( ANY_EVACUATIONS == "Y" | 
                            ANY_INJURIES == "Y" |
                            ANY_FATALITIES == "Y",
                          "H",
                          "NA"),
         dmg = if_else(ANY_DAMAGES == "Y"|
                         FIRE_INVOLVED == "Y",
                       "D",
                       "NA"),
         hd = paste0(human, dmg),
         hd = str_replace(hd, "NANA", "None"),
         hd = str_remove(hd, "NA"))
register_google(key = "AIzaSyBSxdEir9Vz3beRDju9REejDrgJBvA4eng")
#pipes_geocode <- geocode(pipes$LOC_FULL)
#write.csv(pipes_geocode, "pipes_geo.csv")
pipes_geocode <- read_csv("pipes_geo.csv")[,2:3]
pipes <- cbind(pipes, pipes_geocode)

# bw_map <- get_googlemap(center = c(-97.9, 40), zoom = 4,
#   color = "bw",
#   style = "feature:road|visibility:off&style=element:labels|visibility:off&style=feature:administrative|visibility:off")
```




```{r adding other sheets }
pipetype <- read_xlsx(path, sheet = 4) %>%
  select(SEQNOS, PIPELINE_TYPE)

pipes <- left_join(pipes, pipetype, by = "SEQNOS")

mattype <- read_xlsx(path, sheet = 5) %>%
  select(SEQNOS, NAME_OF_MATERIAL)

pipes <- left_join(pipes, mattype, by = "SEQNOS")
```



```{r include = FALSE}
pipes %>%
  count(NAME_OF_MATERIAL)%>%
  group_by(fct_lump(NAME_OF_MATERIAL,n = 4,w = n))%>%
  tally(n)%>%
  rename(NAME_OF_MATERIAL = 1)%>%
  ggplot(aes(x = reorder(NAME_OF_MATERIAL, n), y = n))+
  geom_point(color = blue)+
  geom_segment( aes( xend=reorder(NAME_OF_MATERIAL, n), y=0, yend=n), color = blue)+
 # geom_label(aes(label = n), size = 3, nudge_y = 20)+
  coord_flip()+
  labs(x = "",
       y = "",
       caption = capt,
       title = "NRC Pipeline Calls by Reported Material",
       subtitle = timerange)+
  theme_clean(base_size = 8)+
  theme(plot.title.position = "plot",
        panel.grid.major.x = element_line(linetype = "dotted", colour = "gray35"),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
```


```{r clean data}
cols = c("None" = green, "H" = pink, "D" = yellow, "HD" = orange)

materials <- c("CRUDE", "GAS", "PETROLEUM","KEROSENE","ETHYLENE","DIESEL")


relPipes<- pipes %>%
  filter((hd != "None" | MEDIUM_DESC == "WATER") & 
        grepl(paste(materials, collapse="|"), NAME_OF_MATERIAL, fixed = F)) %>%
  mutate(NUMBER_EVACUATED = replace_na(NUMBER_EVACUATED,0),
         NUMBER_INJURED = replace_na(NUMBER_INJURED, 0),
         NUMBER_FATALITIES = replace_na(NUMBER_FATALITIES, 0),
         PIPELINE_TYPE = str_trunc(PIPELINE_TYPE,8),
         BODY_OF_WATER = replace_na(BODY_OF_WATER, "None"),
         protocol = if_else(grepl("///",DESCRIPTION_OF_INCIDENT),"Y","Z"),
         SEQNOS = str_sub(as.character(SEQNOS), -3))%>%
  arrange(INC_DATE, protocol)%>%
  distinct(LOC_FULL,INC_DATE, .keep_all = T)%>%
  filter(!grepl("ROODHOUSE", LOC_FULL))
```


```{r some reported figures}
fires = filter(count(relPipes, FIRE_INVOLVED), FIRE_INVOLVED == "Y")$n
evac = filter(count(relPipes, ANY_EVACUATIONS), ANY_EVACUATIONS == "Y")$n
evac_n = sum(relPipes$NUMBER_EVACUATED, na.rm = T)
injury = filter(count(relPipes, ANY_INJURIES), ANY_INJURIES == "Y")$n
injury_n = sum(relPipes$NUMBER_INJURED, na.rm = T)
fatal = filter(count(relPipes, ANY_FATALITIES), ANY_FATALITIES == "Y")$n
fatal = paste(fatal)
fatal_n = sum(relPipes$NUMBER_FATALITIES, na.rm = T)
damage = filter(count(relPipes, ANY_DAMAGES), ANY_DAMAGES == "Y")$n
offshore = filter(count(relPipes, OFFSHORE), OFFSHORE == "Y")$n
states = unique(relPipes$LOCATION_STATE)
states_n = length(states)

phmsa_r = nrow(filter(relPipes, grepl("PHMSA", FEDERAL_AGENCY_NOTIFIED)))

```

Calls counted here include those which reported spills of `r str_to_lower(materials)` and either caused direct human impact (injury, evacuation, or fatality), direct damage (fire, reported damage), or occured in water. `r nrow(relPipes)` calls fitting these criteria occurred over the past 2 weeks in `r states_n` states : `r states`. There may be some discrepancies between the list and the table, as there are occasionally instances where a column like "ANY_EVACUATION" returns "Y" (yes), but the number is either 0 or unknown at the time of the call or update.  

Number of calls that reported:  
  
* Fire: `r fires`  
* Evacuation: `r evac` (`r evac_n` evacuated)  
* Injury: `r injury` (`r injury_n` injured)  
* Fatality: `r fatal` (`r fatal_n` dead)  
* Damage: `r damage`  
* Offshore Spill: `r offshore`  
* to PHMSA: `r phmsa_r`

```{r include = TRUE}

relPipes %>%
  select(LOC_FULL, INC_DATE,INCIDENT_CAUSE,
         PIPELINE_TYPE,NAME_OF_MATERIAL,FIRE_INVOLVED,
         NUMBER_EVACUATED,NUMBER_INJURED,NUMBER_FATALITIES,
         BODY_OF_WATER,protocol )%>%
  mutate(LOC_FULL = str_wrap(LOC_FULL, 8),
         NAME_OF_MATERIAL= if_else(NAME_OF_MATERIAL == "NATURAL GAS",
                                   "NG","HL"),
         INCIDENT_CAUSE = str_trunc(INCIDENT_CAUSE, 10),
         protocol = if_else(protocol=="Y", T,F),
         INC_DATE = paste0(month(INC_DATE),"-",day(INC_DATE)))%>%
  kbl(col.names = c("Place","Date","Cause","Type","Material",
                    "Fire?","Evac.","Inj.","Ftl.", "Water", "Updated?"))%>%
  kable_paper("striped", font_size = 8, full_width = F)
```

```{r map setup}
state.file <- "/Users/jameseager/Library/CloudStorage/OneDrive-pstrust.org/DataViz/GIS/Pipelines/Processing/us_states.shp"

projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

states.proj<-read_sf(state.file)

#mb_access_token("pk.eyJ1IjoiamltZWFnYWgiLCJhIjoiY2l2aWw3YmJnMDIxdzJ5bXFuZDZ4bnU1YiJ9.lMwViZI1fDl75hf6XGm78g", install = T)

sat_tiles <- get_static_tiles(location = states.proj,
                              zoom = 4,
                              style_id = "light-v10",
                              username = "mapbox",
                              crop = T)

spPipes <- st_as_sf(filter(relPipes, !is.na(lon)), 
                    coords = c("lon", "lat"), 
                    crs = projcrs)

```

```{r include = TRUE, fig.fullwidth = TRUE}
cols = c("None" = green, "H" = pink, "D" = yellow, "HD" = orange)


spPipes %>%
  ggplot()+
  layer_spatial(sat_tiles)+
  geom_sf(data = states.proj, fill = NA, color = "gray25", size = .5)+
  geom_sf(aes(color = hd, shape = MEDIUM_DESC), alpha = .75)+
  scale_shape(name = "Release Medium")+
  scale_color_manual(values = cols,
                     labels = c("Neither","Human","Damage","Both"),
                     name = "Impact:")+
  coord_sf(crs = 3857,
           xlim = c(-13692297.369799, -6901808.430296),
           ylim = c(2511525.230043, 6446275.839285))+
  theme_void()+
  theme(legend.text = element_text(size = 3),
        legend.title = element_text(size = 4),
        legend.spacing = unit(.3,"cm"))+
  labs(caption = capt,
       title = "Pipeline Incident Call Locations")
    
## SUPERSEDED
# ggmap(bw_map, extent = "panel")+
#   geom_path(data = states.proj, aes(long, lat, group = group), colour = "gray15")+
#   geom_point(data = relPipes, aes(x = lon, y = lat, color = hd, shape = MEDIUM_DESC),
#              alpha = .8)+
#   geom_text_repel(data = relPipes, aes(x = lon, y = lat, label = SEQNOS), 
#                   nudge_x = 0.05, size = 2, min.segment.length = 0)+
#   scale_color_manual(values = cols,
#                      labels = c("Neither","Human","Damage","Both"),
#                      name = "Impact:")+
#   theme_void()+
#   theme(legend.position = "right",
#         legend.justification = "center",
#         legend.direction = "vertical",
#         legend.background = element_blank(),
#         legend.key = element_blank(),
#         legend.text = element_text(size = 3),
#         legend.title = element_blank())+
#   labs(caption = capt,
#        title = "Pipeline Incident Calls to NRC",
#        subtitle = timerange)
```

