---
title: "NRC Rundown (2 weeks)"
subtitle: "Pipeline Incident Calls (2 weeks)"
runningheader: "Pipeline Safety Trust `r format(Sys.time(), '%m/%Y')`"
author: "Pipeline Safety Trust"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_handout: default
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{img/PST_logo.png}\LARGE\\}
  - \posttitle{\end{center}}
editor_options: 
  chunk_output_type: console
---

```{r delete all, include = FALSE, echo = FALSE}
rm(list = ls())
```

```{r setup, include=FALSE}
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

#relevant libs, some more useful than others
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(ggthemes)
library(scales)
library(readxl)
library(kableExtra)
library(ggmap)
library(sf)
library(tufte)
library(ggspatial)
library(mapboxapi)

#geocoding of locations
source("nrc_keys.R")

#some general knit options
knitr::opts_chunk$set(echo = FALSE, include = FALSE, message = FALSE, warning = FALSE,
                      kable.force.latex = T)
options(knitr.table.format = "latex")
```

```{r color setup}
colors <- brewer.pal(6,"Set2")
green <- colors[1]
orange <- colors[2]
blue <- colors[3]
pink <- colors[4]
limegreen <- colors[5]
yellow <- colors[6]

#pst 
darkBlue <- "#003E59"
midBlue <- "#00546F"
bluGreen <- "#459197"
liteGreen <- "#61A893"
```

```{r theme_pst}
library(showtext)
#> Loading required package: sysfonts
#> Loading required package: showtextdb
pstFont = "Montserrat"
font_add_google(pstFont)
showtext_auto()
theme_pst <- function(baseSize=8) {
  (theme(
                       axis.line.x = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.line.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       axis.ticks.y = element_line(
                         colour = "#182125",
                         size = 0.5,
                         linetype = "solid"
                       ),
                       text = element_text(family = pstFont),
                       axis.text = element_text(colour = "#8C9394",
                                                size = baseSize * .8),
                       axis.title = element_text(colour ="black"),
                       panel.grid.minor.y = element_line(linetype = "dotted", 
                                                         colour = "#C4C8C6"),
                       panel.grid.major.y = element_line(colour = "#394C56", 
                                                         linetype = "dotted"),
                       panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(),
                       panel.background = element_blank(),
                       panel.border = element_blank(),
                       strip.background = element_rect(linetype = 0),
                       strip.text = element_blank(),
                       strip.text.x = element_blank(),
                       strip.text.y = element_blank(),
                       legend.text = element_text(colour = "black",
                                                  size = baseSize * .8),
                       legend.background = element_rect(fill = NULL, 
                                                        color = "#182125"),
                       legend.title = element_text( face = "bold", 
                                                    colour = "black",
                                                    size = baseSize),
                       legend.position = "right",
                       legend.key = element_blank(),
                       #legend.background = element_blank(),
                       plot.tag.position = "topright",
                       plot.tag = element_text(size = baseSize *.66,
                                               colour = "#8C9394"),
                       plot.background = element_blank(),
                       plot.title = element_text(face = "bold", 
                                                 colour = "black",
                                                 size = baseSize*1.2),
                       plot.subtitle = element_text(colour = "#8C9394",
                                                    size = baseSize)
                     )
  )
}
```

```{r}

pipes <- read_csv("pipes_geo.csv") %>%
  left_join( read_xlsx(temp, sheet = 3), by = "SEQNOS", suffix = c("",".x"))%>%    #incident details
  left_join( read_xlsx(temp, sheet = 4), by = "SEQNOS", suffix = c("",".y"))%>%    #incident details
  left_join( read_xlsx(temp, sheet = 5), by = "SEQNOS", suffix = c("",".z"))      #material details

#set another text bit for plotting
tr <- stamp("03.22.87")
timerange = paste("From", 
                  tr((max(pipes$INC_DATE) - weeks(2))),
                  "to",
                  tr(max(pipes$INC_DATE)))
#generic caption text, update to reflect data 
cp <- stamp("8 March, 2022")
capt = paste("Data from NRC FOIA:", timerange)
```





```{r include = FALSE}
pipes %>%
  count(NAME_OF_MATERIAL)%>%
  group_by(fct_lump(NAME_OF_MATERIAL,n = 4,w = n))%>%
  tally(n)%>%
  rename(NAME_OF_MATERIAL = 1)%>%
  ggplot(aes(x = reorder(NAME_OF_MATERIAL, n), y = n))+
  geom_point(color = darkBlue)+
  geom_segment( aes( xend=reorder(NAME_OF_MATERIAL, n), y=0, yend=n), color = blue)+
 # geom_label(aes(label = n), size = 3, nudge_y = 20)+
  coord_flip()+
  labs(x = "",
       y = "",
       caption = capt,
       title = "NRC Pipeline Calls by Reported Material",
       subtitle = timerange)+
  theme_pst(baseSize = 12)+
  theme(plot.title.position = "plot",
        panel.grid.major.x = element_line(linetype = "dotted", colour = "gray35"),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
```


```{r clean data}
cols = c("None" = green, "H" = pink, "D" = yellow, "HD" = orange)

materials <- c("CRUDE", "GAS", "PETROLEUM","KEROSENE","ETHYLENE","DIESEL")


relPipes<- pipes %>%
  #mutate(sig = if_else())
  filter((hd != "None" | MEDIUM_DESC == "WATER") & 
        grepl(paste(materials, collapse="|"), NAME_OF_MATERIAL, fixed = F)) %>%
  mutate(NUMBER_EVACUATED = replace_na(NUMBER_EVACUATED,0),
         NUMBER_INJURED = replace_na(NUMBER_INJURED, 0),
         NUMBER_FATALITIES = replace_na(NUMBER_FATALITIES, 0),
         PIPELINE_TYPE = str_trunc(PIPELINE_TYPE,8),
         BODY_OF_WATER = replace_na(BODY_OF_WATER, "None"),
         protocol = if_else(grepl("///",DESCRIPTION_OF_INCIDENT),"Y","Z"),
         SEQNOS = str_sub(as.character(SEQNOS), -3))%>%
  arrange(INC_DATE, protocol)%>%
  distinct(LOC_FULL,INC_DATE, .keep_all = T)%>%
  filter(!grepl("ROODHOUSE", LOC_FULL))
```


```{r some reported figures}
fires = filter(count(relPipes, FIRE_INVOLVED), FIRE_INVOLVED == "Y")$n
evac = filter(count(relPipes, ANY_EVACUATIONS), ANY_EVACUATIONS == "Y")$n
evac_n = sum(relPipes$NUMBER_EVACUATED, na.rm = T)
injury = filter(count(relPipes, ANY_INJURIES), ANY_INJURIES == "Y")$n
injury_n = sum(relPipes$NUMBER_INJURED, na.rm = T)
fatal = filter(count(relPipes, ANY_FATALITIES), ANY_FATALITIES == "Y")$n
fatal = paste(fatal)
fatal_n = sum(relPipes$NUMBER_FATALITIES, na.rm = T)
damage = filter(count(relPipes, ANY_DAMAGES), ANY_DAMAGES == "Y")$n
offshore = filter(count(relPipes, OFFSHORE), OFFSHORE == "Y")$n
states = unique(relPipes$LOCATION_STATE)
states_n = length(states)

phmsa_r = nrow(filter(relPipes, grepl("PHMSA", FEDERAL_AGENCY_NOTIFIED)))

```

Calls counted here include those which reported spills of `r str_to_lower(materials)` and either caused direct human impact (injury, evacuation, or fatality), direct damage (fire, reported damage), or occured in water. `r nrow(relPipes)` calls fitting these criteria occurred over the past 2 weeks in `r states_n` states : `r states`. There may be some discrepancies between the list and the table, as there are occasionally instances where a column like "ANY_EVACUATION" returns "Y" (yes), but the number is either 0 or unknown at the time of the call or update.  

Number of calls that reported:  
  
* Fire: `r fires`  
* Evacuation: `r evac` (`r evac_n` evacuated)  
* Injury: `r injury` (`r injury_n` injured)  
* Fatality: `r fatal` (`r fatal_n` dead)  
* Damage: `r damage`  
* Offshore Spill: `r offshore`  
* to PHMSA: `r phmsa_r`

```{r include = TRUE}

relPipes %>%
  select(LOC_FULL, INC_DATE,INCIDENT_CAUSE,
         PIPELINE_TYPE,NAME_OF_MATERIAL,FIRE_INVOLVED,
         NUMBER_EVACUATED,NUMBER_INJURED,NUMBER_FATALITIES,
         BODY_OF_WATER,protocol )%>%
  mutate(LOC_FULL = str_wrap(LOC_FULL, 8),
         NAME_OF_MATERIAL= if_else(NAME_OF_MATERIAL == "NATURAL GAS",
                                   "NG","HL"),
         INCIDENT_CAUSE = str_trunc(INCIDENT_CAUSE, 10),
         protocol = if_else(protocol=="Y", T,F),
         INC_DATE = paste0(month(INC_DATE),"-",day(INC_DATE)))%>%
  kbl(col.names = c("Place","Date","Cause","Type","Material",
                    "Fire?","Evac.","Inj.","Ftl.", "Water", "Updated?"))%>%
  kable_paper("striped", font_size = 8, full_width = F)
```

```{r map setup}
state.file <- "/Users/jameseager/Library/CloudStorage/OneDrive-pstrust.org/DataViz/GIS/Pipelines/Processing/us_states.shp"

projcrs <- "+proj=longlat +datum=WGS84 +no_defs"

states.proj<-read_sf(state.file)

#mb_access_token("pk.eyJ1IjoiamltZWFnYWgiLCJhIjoiY2l2aWw3YmJnMDIxdzJ5bXFuZDZ4bnU1YiJ9.lMwViZI1fDl75hf6XGm78g", install = T)

sat_tiles <- get_static_tiles(location = states.proj,
                              zoom = 4,
                              style_id = "light-v10",
                              username = "mapbox",
                              crop = T)

spPipes <- st_as_sf(filter(relPipes, !is.na(lon)), 
                    coords = c("lon", "lat"), 
                    crs = projcrs)

```

```{r include = TRUE, fig.fullwidth = TRUE}
cols = c("None" = green, "H" = pink, "D" = yellow, "HD" = orange)


spPipes %>%
  ggplot()+
  layer_spatial(sat_tiles)+
  geom_sf(data = states.proj, fill = NA, color = "gray25", size = .5)+
  geom_sf(aes(color = hd, shape = MEDIUM_DESC), alpha = .75)+
  scale_shape(name = "Release Medium")+
  scale_color_manual(values = cols,
                     labels = c("Neither","Human","Damage","Both"),
                     name = "Impact:")+
  coord_sf(crs = 3857,
           xlim = c(-13692297.369799, -6901808.430296),
           ylim = c(2511525.230043, 6546275.839285))+
  theme_void()+
  theme(legend.text = element_text(size = 3),
        legend.title = element_text(size = 4),
        legend.spacing = unit(.3,"cm"),
        legend.position = c(0.9,0.05))+
  labs(caption = paste0(capt, "\n (c) Mapbox, (c) OpenStreetMap"),
       title = "Pipeline Incident Call Locations")
 
```

